---
title: "Rsa-uid-testing"
author: "Ben"
date: "1/22/2018"
output: html_document
---

# Desiderata for rsa-uid model

### (1) Preferential marking of high-surprisal utterances by speakers
The basic UID finding is that speakers are more likely to mark (add a complementizer, use a full form instead of contraction, etc.) items that are high surprisal.

* We should expect this preference to increase as a function of surprisal for noise and non-noise models.
* We should expect this preference to increase as a function of theta for noise models.

### (2) Um-implicature for listeners
Listeners who are aware of the production process should infer that the content following a marked term (or consisting just of a marked term) should be followed by a high-surprisal item.

* We should expect this preference to increase as a function of surprisal for noise and non-noise models.
* We should expect this preference to increase as a function of theta for noise models.

### (2) UID-corpus effect.
A corpus of utterances generate by a UID-speaker should display the standard relationship described in Levy & Jaeger (2007) -- the probability of a marked utterance should increase with the suprisal of the following content under a language model

* We should expect this finding to be modulated by theta, but robust across different levels of theta.

# Tests

Things to normalize across models:

* Model parameters
  + input
  + theta
  + utterance set: ['X a', 'a', 'X b', 'b', 'X c', 'c', 'X c', 'c']
  + utterance probs: [marked: 0.5, unmarked: 1]
  + worlds: [A, B, C, D] from high --> low surprisal
  + world probs: chosen to approximate range of surprisal values in Levy & Jaeger (2007)
  + recursion level

```{r warning=FALSE, message=FALSE}
library(doParallel)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(tidyr)
setwd('/Users/benpeloquin/Desktop/Projects/rsa_uid/')
```

```{r helpers}
getModelFile <- function(fPath) {
  readChar(fPath, file.info(fPath)$size)
}

create_test <- function(fPath, run_fn, test_fn) {
  m <- getModelFile(fPath)
  d <- run_fn(m)
  return(function() test_fn(d))
}

THETA <- 0.1
```

## Noisy RSA1
### Listener test
```{r test_1}
listener_run_fn_ <- function(modelFile, alpha=1, lambda=10, modelName='L1', theta=0.1) {
  rData <- data.frame(alpha=alpha, lambda=lambda, modelName=modelName, theta=theta)
  rwebppl::webppl(modelFile, data=rData, data_var='rData')
}
listener_test_fn_ <- function(df) {
  filter(df, d=='a')$prob > filter(df, d=='b')$prob
}
```

### Speaker test
```{r test_2}
speaker_run_fn_ <- function(modelFile, alpha=5, lambda=1, modelName='S3', theta=theta, speakerInput='a'){
  noisy_rsa1_speaker_run_fn_ <- function(modelFile, input=speakerInput) {
    # Run Settings
    # ------------
    rData <- data.frame(alpha=alpha, lambda=lambda, modelName=modelName, theta=theta, speakerInput=input)
    rwebppl::webppl(modelFile, data=rData, data_var='rData')
  }
  
  d1 <- noisy_rsa1_speaker_run_fn_(modelFile, 'a') %>%
    mutate(input='a')
  d2 <- noisy_rsa1_speaker_run_fn_(modelFile, 'b') %>%
    mutate(input='b')
  d_full <- rbind(d1, d2)
  d_full
}

speaker_test_fn_ <- function(df) {
  # Test one, check proportions
  dfa <- filter(df, input=='a')
  dfb <- filter(df, input=='b')
  propAMarked <- filter(dfa, support=='X a')$prob / (filter(dfa, support=='X a')$prob + filter(dfa, support=='a')$prob)
  propBMarked <- filter(dfb, support=='X b')$prob / (filter(dfb, support=='X b')$prob + filter(dfb, support=='b')$prob)
  propAMarked > propBMarked
}

speaker_test_fn_ <- function(df) {
  # Test one, check proportions
  filter(df, input=='a' & support == 'X a')$prob > filter(df, input=='b' & support == 'X b')$prob
}
```

## Corpus test
```{r corpus_run_helpers}
createRunFn <- function(modelStr) {
    return(function(data, dataName='rData') { 
      rwebppl::webppl(modelStr, data=data, data_var=dataName)
    })
}

createRData_ <- function(modelName, alpha, lambda, theta, n) {
  data.frame(modelName=modelName,
             alpha=alpha,
             lambda=lambda,
             theta=theta,
             n=n,
             stringsAsFactors=FALSE)
}

# Pipeline
# ---------
# (1) run experiments in parallel
# (2) run experiments in parallel

runExperimentFn_ <- function(runFn, modelName, alpha, lambda, theta, n) {
  return(function(expNum) {
    rData <- createRData_(modelName, alpha, lambda, theta, n)
    df <- runFn(rData) %>%
      mutate(modelName=modelName,
             alpha=alpha,
             lambda=lambda,
             theta=theta,
             n=n,
             expNum=expNum)
    df
  })
}

createLM_ <- function(dfCorpus) {
  annotateUtterance <- function(x) {
    x <- gsub(' ', '', x)
    x <- paste0('^', x)
    x <- paste0(x, '$')
    return(x)
  }
  stream <- dfCorpus %>%
          mutate(annotatedUtterance=annotateUtterance(currUtterance)) %>%
          select(annotatedUtterance) %>%
          unlist %>%
          paste0(collapse='') %>%
          gsub(' ', '', ., fixed=TRUE)
  ng <- ngram::ngram(stream, n=2, sep='')  # bigram model
  dfNgramRaw <- data.frame(ngram::get.phrasetable(ng))
  dfNgramRaw$first <- sapply(dfNgramRaw$ngrams, function(x) {strsplit(x, '')}[[1]][1])
  dfNgramRaw$second <- sapply(dfNgramRaw$ngrams, function(x) {strsplit(x, '')}[[1]][3])
  dfNgramRaw <- dfNgramRaw %>%
    mutate(first=ifelse(first==' ', '*', first),
           second=ifelse(second==' ', '*', second))
  dfNgramRaw$theta <- unique(dfCorpus$theta)
  dfNgramRaw$alpha <- unique(dfCorpus$alpha)
  dfNgramRaw$lambda <- unique(dfCorpus$lambda)
  dfNgramRaw$expNum <- unique(dfCorpus$expNum)
  dfNgramRaw
}

run_corpus_analysis_ <- function(df_lm) {
  df_lm %>%
    group_by(expNum, second) %>%
    mutate(secondTotalProp=sum(prop),       # p(a|X) + p(a|^)
           lik=prop/secondTotalProp) %>%    # p(X|a) = p(X a) / p(a|X) + p(a|^)
    ungroup %>%
    group_by(expNum) %>%
    mutate(postProb=log2(secondTotalProp/sum(prop))) %>%  # p(a)
    ungroup %>%
    select(expNum, first, second, lik, postProb) %>%
    filter(second %in% c(letters[1:5], '*'), first=='X') %>%
    group_by(second) %>%
    summarise(
      avgPostProb=mean(postProb),
      xmin=quantile(postProb, probs=c(0.05)),
      xmax=quantile(postProb, probs=c(0.95)),
      avgLikelihood=mean(lik),
      ymin=quantile(lik, probs=c(0.05)),
      ymax=quantile(lik, probs=c(0.95)),
      n=n())
}

process_sims_ <- function(sims) {
  lSimsExp1 <- sims %>%
    group_by(expNum) %>%
    do(vals=data.frame(.)) %>%
    lapply(function(x) {(x)})

  lD <- lapply(lSimsExp1$vals, function(x) {createLM_(x)})
  df <- do.call(rbind, lD)
  df
}



# Put in a function somewhere
# no_cores <- detectCores() - 1
# cl <- makeCluster(no_cores, type='FORK')
# registerDoParallel(cl)
# 
# simsExp1 <- foreach(i=seq(1, nIters), .packages=c('dplyr', 'rwebppl'), .combine=rbind) %dopar% runExp1(i)
# stopCluster(cl)
```

```{r corpus_run_fn}
# corpus_run_fn
# -------------
# Pipeline
# (1) run simulations
# (2) create lm for each simluation
# (3) run lm analysis per simulation - utterance surprisal X likelihood of marked
corpus_run_fn_ <- function(modelFile, modelName='S3', alpha=5, lambda=1, theta=0.2, nUtterances=1000){
  runFn <- createRunFn(modelFile)
  runExp_ <- runExperimentFn_(runFn, modelName, alpha, lambda, theta, nUtterances)
  # Register cores
  no_cores <- detectCores() - 1
  cl <- makeCluster(no_cores, type='FORK')
  registerDoParallel(cl)
  # Run sims
  ptm <- proc.time()
  nSims <- 100
  cat("Running", nSims, "simulations each with", nUtterances, "samples.")
  sims <- foreach(i=seq(1, nSims), .packages=c('dplyr', 'rwebppl'), .combine=rbind) %dopar% runExp_(i)
  stopCluster(cl)
  etm <- proc.time() - ptm
  cat("runtime: ", etm[3] / 60)
  # Create LMs
  df_lm <- process_sims_(sims)
  # LM analysis
  df_summary <- run_corpus_analysis_(df_lm)
  df_summary
}

corpus_test_fn_ <- function(df) {
  return(
    (filter(df, second=='a')$avgPostProb < filter(df, second=='b')$avgPostProb) && 
    (filter(df, second=='a')$avgLikelihood > filter(df, second=='b')$avgLikelihood))
}
```

```{r run_tests_helper}
run_all_tests <- function(tests){
  sapply(tests, function(x) x())
}
```

# Tests

## Basic RSA - no noise
```{r run_basic_rsa1_tests}
basic_rsa1_listener_file <- 'Models/testing/basic-rsa1-listener.wppl'
basic_rsa1_listener_test <- create_test(basic_rsa1_listener_file, listener_run_fn_, listener_test_fn_)
basic_rsa1_speaker_file <- 'Models/testing/basic-rsa1-speaker.wppl'
basic_rsa1_speaker_test <- create_test(basic_rsa1_speaker_file, speaker_run_fn_, speaker_test_fn_)
basic_rsa1_corpus_file <- 'Models/testing/basic-rsa1-corpus.wppl'
basic_rsa1_corpus_test <- create_test(basic_rsa1_corpus_file, corpus_run_fn_, corpus_test_fn_)
basic_rsa1_tests <- c(basic_rsa1_listener_test, basic_rsa1_speaker_test, basic_rsa1_corpus_test)
run_all_tests(basic_rsa1_tests)
```

```{r noisy_rsa3_corpus_plot}
basic_rsa1_corpus_file <- 'Models/testing/basic-rsa1-corpus.wppl'
df_brsas1 <- corpus_run_fn_(getModelFile(basic_rsa1_corpus_file))
df_brsas1 %>% 
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=as.factor(second))) +
    geom_point() +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin)) +
    ylab('Proportion marked') +
    xlab('Average surprisal') +
    scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, length.out=6)) +
    theme_few()
```

## Noisy RSA models
```{r run_noisy_rsa1_tests}
# Static theta, S3 only samples all utterances
noisy_rsa1_listener_file <- 'Models/testing/noisy-rsa1-listener.wppl'
noisy_rsa1_listener_test <- create_test(noisy_rsa1_listener_file, listener_run_fn_, listener_test_fn_)
noisy_rsa1_speaker_file <- 'Models/testing/noisy-rsa1-speaker.wppl'
noisy_rsa1_speaker_test <- create_test(noisy_rsa1_speaker_file, speaker_run_fn_, speaker_test_fn_)
noisy_rsa1_corpus_file <- 'Models/testing/noisy-rsa1-corpus.wppl'
noisy_rsa1_corpus_test <- create_test(noisy_rsa1_corpus_file, corpus_run_fn_, corpus_test_fn_)
noisy_rsa1_tests <- c(noisy_rsa1_listener_test, noisy_rsa1_speaker_test, noisy_rsa1_corpus_test)
run_all_tests(noisy_rsa1_tests)
```


```{r noisy_rsa1_corpus_plot}
noisy_rsa1_corpus_file <- 'Models/testing/noisy-rsa1-corpus.wppl'
df_nrsa1 <- corpus_run_fn_(getModelFile(noisy_rsa1_corpus_file), modelName='S1')
df_nrsa1 %>% 
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=as.factor(second))) +
    geom_point() +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin), height=0) +
    ylab('Likelihood of marked version') +
    xlab('mean(log(p(utt))') +
    scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, length.out=6)) +
    theme_few()
```

```{r noisy_rsa1_vary_spaker_level}
noisy_rsa1_corpus_file <- 'Models/testing/noisy-rsa1-corpus.wppl'
models <- c('S1', 'S2', 'S3', 'S4', 'S5')
noisy_rsa1_dfs <- data.frame()
for (model in models) {
  cat("Running model:", model)
  df_curr_run <- corpus_run_fn_(getModelFile(noisy_rsa1_corpus_file), modelName=model) %>%
    mutate(model=model)
  noisy_rsa1_dfs <- rbind(noisy_rsa1_dfs, df_curr_run)
}
```

```{r noisy_rsa1_vary_spaker_level_plot}
noisy_rsa1_dfs %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=model)) +
    geom_point() +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
    ylab('Proportion marked') +
    xlab('Average surprisal') +
    scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, length.out=6)) +
    theme_few()
```


```{r run_noisy_rsa1_tests}
# Static theta, S3 only samples valid utterances.
noisy_rsa2_listener_file <- 'Models/testing/noisy-rsa2-listener.wppl'
noisy_rsa2_listener_test <- create_test(noisy_rsa2_listener_file, listener_run_fn_, listener_test_fn_)
noisy_rsa2_speaker_file <- 'Models/testing/noisy-rsa2-speaker.wppl'
noisy_rsa2_speaker_test <- create_test(noisy_rsa2_speaker_file, speaker_run_fn_, speaker_test_fn_)
noisy_rsa2_corpus_file <- 'Models/testing/noisy-rsa2-corpus.wppl'
noisy_rsa2_corpus_test <- create_test(noisy_rsa2_corpus_file, corpus_run_fn_, corpus_test_fn_)
noisy_rsa2_tests <- c(noisy_rsa2_listener_test, noisy_rsa2_speaker_test, noisy_rsa2_corpus_test)
run_all_tests(noisy_rsa2_tests)
```

```{r noisy_rsa2_corpus_plot}
noisy_rsa2_corpus_file <- 'Models/testing/noisy-rsa2-corpus.wppl'
df_nrsa2 <- corpus_run_fn_(getModelFile(noisy_rsa2_corpus_file))
df_nrsa2 %>% 
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=as.factor(second))) +
    geom_point() +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
    ylab('Proportion marked') +
    xlab('Average surprisal') +
    scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, length.out=6)) +
    theme_few()
```

```{r compare}
df_brsas1 <- df_brsas1 %>%
  mutate(model='brsas1')
df_nrsa1 <- df_nrsa1 %>%
  mutate(model='nrsa1')
df_nrsa2 <- df_nrsa2 %>%
  mutate(model='nrsa2')
df_nrsa3 <- df_nrsa3 %>%
  mutate(model='nrsa3')
df_all <- rbind(df_brsas1, df_nrsa1, df_nrsa2, df_nrsa3)
```

```{r compare_plot}
df_all %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=model)) +
    geom_point() +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
    ylab('Proportion marked') +
    xlab('Average surprisal') +
    scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, length.out=6)) +
    theme_few()
```

```{r run_noisy_rsa1_tests}
# Lifted theta - spekaers / listeners infer noise level...
noisy_rsa3_listener_file <- 'Models/testing/noisy-rsa3-listener.wppl'
noisy_rsa3_listener_test <- create_test(noisy_rsa3_listener_file, listener_run_fn_, listener_test_fn_)
noisy_rsa3_speaker_file <- 'Models/testing/noisy-rsa3-speaker.wppl'
noisy_rsa3_speaker_test <- create_test(noisy_rsa3_speaker_file, speaker_run_fn_, speaker_test_fn_)
noisy_rsa3_corpus_file <- 'Models/testing/noisy-rsa3-corpus.wppl'
noisy_rsa3_corpus_test <- create_test(noisy_rsa3_corpus_file, corpus_run_fn_, corpus_test_fn_)
noisy_rsa3_tests <- c(noisy_rsa3_listener_test, noisy_rsa3_speaker_test, noisy_rsa3_corpus_test)
run_all_tests(noisy_rsa3_tests)
```

```{r noisy_rsa3_corpus_plot}
noisy_rsa3_corpus_file <- 'Models/testing/noisy-rsa3-corpus.wppl'
df_nrsa3 <- corpus_run_fn_(getModelFile(noisy_rsa3_corpus_file))
df_nrsa3 %>% 
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=as.factor(second))) +
    geom_point() +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
    ylab('Proportion marked') +
    xlab('Average surprisal') +
    scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, length.out=6)) +
    theme_few()
```

# Explore

## Effect of theta on slopes

#### noisy-rsa with constrained utterances
```{r noisy_rsa2_thetas}
thetas <- c(0.1, 0.9)
df_noisy_rsa2 <- data.frame()
for (theta in thetas) {
  cat("Runing theta:", theta)
  df_curr_run <- corpus_run_fn_(getModelFile(noisy_rsa2_corpus_file), theta=theta) %>%
    mutate(theta=theta)
  df_noisy_rsa2 <- rbind(df, df_curr_run)
}
```

```{r noisy_rsa1_thetas_plot}
df_noisy_rsa2 %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=as.factor(theta))) +
    geom_point() +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
    ylab('Proportion marked') +
    xlab('Average surprisal') +
    scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, length.out=6)) +
    theme_few()
```

#### noisy-rsa without constrained utterances
```{r noisy_rsa1_thetas}
thetas <- c(0.1, 0.3, 0.5, 0.7, 0.9)
df_noisy_rsa1 <- data.frame()
for (theta in thetas) {
  cat("Runing theta:", theta)
  df_curr_run <- corpus_run_fn_(getModelFile(noisy_rsa1_corpus_file), theta=theta) %>%
    mutate(theta=theta)
  df_noisy_rsa1 <- rbind(df, df_curr_run)
}
```

```{r noisy_rsa1_thetas_plot}
df_noisy_rsa1 %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=as.factor(theta))) +
    geom_point() +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
    ylab('Proportion marked') +
    xlab('Average surprisal') +
    scale_y_continuous(limits=c(0, 1), breaks=seq(0, 1, length.out=6)) +
    theme_few()
```
Sample from all utterances, vary theta, S3 speaker.


```{r noisy_rsa1_listeners}
noisy_rsa1_listener_file <- 'Models/testing/noisy-rsa1-listener.wppl'
df_noisy_rsa1_listeners <- data.frame()
for (theta in seq(0.1, 0.9, by=0.1)) {
  df_curr <- listener_run_fn_(getModelFile(noisy_rsa1_listener_file), theta=theta) %>%
    mutate(theta=theta)
  df_noisy_rsa1_listeners <- rbind(df_noisy_rsa1_listeners, df_curr)  
}
```

```{r noisy_rsa1_listeners_plot}
df_noisy_rsa1_listeners %>%
  ggplot(aes(x=d, y=prob)) +
    geom_bar(stat='identity') +
    facet_wrap(~theta)
```

```{r noisy_rsa1_speakers}
noisy_rsa1_speaker_file <- 'Models/testing/noisy-rsa1-speaker.wppl'
df_noisy_rsa1_speakers <- data.frame()
for (theta in seq(0.1, 0.9, by=0.1)) {
  df_curr <- speaker_run_fn_(getModelFile(noisy_rsa1_speaker_file), theta=theta) %>%
    mutate(theta=theta)
  df_noisy_rsa1_speakers <- rbind(df_noisy_rsa1_speakers, df_curr)  
}
```

```{r noisy_rsa1_speakers_plot}
df_noisy_rsa1_speakers %>%
  ggplot(aes(x=theta, y=prob)) +
    geom_line(stat='identity') +
    facet_grid(input~support) +
    theme_few()
```

