---
title: "Rsa-uid-testing"
author: "Ben"
date: "1/22/2018"
output: html_document
---

# Desiderata for rsa-uid model

### (1) Preferential marking of high-surprisal utterances by speakers
The basic UID finding is that speakers are more likely to mark (add a complementizer, use a full form instead of contraction, etc.) items that are high surprisal.

* We should expect this preference to increase as a function of surprisal for noise and non-noise models.
* We should expect this preference to increase as a function of theta for noise models.

### (2) Um-implicature for listeners
Listeners who are aware of the production process should infer that the content following a marked term (or consisting just of a marked term) should be followed by a high-surprisal item.

* We should expect this preference to increase as a function of surprisal for noise and non-noise models.
* We should expect this preference to increase as a function of theta for noise models.

### (2) UID-corpus effect.
A corpus of utterances generate by a UID-speaker should display the standard relationship described in Levy & Jaeger (2007) -- the probability of a marked utterance should increase with the suprisal of the following content under a language model

* We should expect this finding to be modulated by theta, but robust across different levels of theta.

# Tests

Things to normalize across models:

* Model parameters
  + input
  + theta
  + utterance set
  + utterance probs
  + worlds
  + world probs
  + recursion level

```{r warning=FALSE, message=FALSE}
library(doParallel)
library(dplyr)
library(ggplot2)
library(tidyr)
setwd('/Users/benpeloquin/Desktop/Projects/rsa_uid/')
```

```{r helpers}
getModelFile <- function(fPath) {
  readChar(fPath, file.info(fPath)$size)
}

create_test <- function(fPath, run_fn, test_fn) {
  m <- getModelFile(fPath)
  d <- run_fn(m)
  return(function() test_fn(d))
}

ALPHA <- 5
LAMBDA <- 50
THETA <- 0.9
```

## Noisy RSA1
### Listener test
```{r test_1}
listener_run_fn_ <- function(modelFile) {
  rData <- data.frame(alpha=ALPHA, lambda=LAMBDA, modelName='L1', theta=THETA)
  rwebppl::webppl(modelFile, data=rData, data_var='rData')
}
listener_test_fn_ <- function(df) {
  filter(df, d=='a')$prob > filter(df, d=='b')$prob
}
```

### Speaker test
```{r test_2}
speaker_run_fn_ <- function(modelFile){
  noisy_rsa1_speaker_run_fn_ <- function(modelFile, input) {
    rData <- data.frame(alpha=ALPHA, lambda=LAMBDA, modelName='S1', theta=THETA, speakerInput=input)
    rwebppl::webppl(modelFile, data=rData, data_var='rData')
  }
  
  d1 <- noisy_rsa1_speaker_run_fn_(modelFile, 'a') %>%
    mutate(input='a')
  d2 <- noisy_rsa1_speaker_run_fn_(modelFile, 'b') %>%
    mutate(input='b')
  d_full <- rbind(d1, d2)
  d_full
}

speaker_test_fn_ <- function(df) {
  # Test one, check proportions
  dfa <- filter(df, input=='a')
  dfb <- filter(df, input=='b')
  propAMarked <- filter(dfa, support=='X a')$prob / (filter(dfa, support=='X a')$prob + filter(dfa, support=='a')$prob)
  propBMarked <- filter(dfb, support=='X b')$prob / (filter(dfb, support=='X b')$prob + filter(dfb, support=='b')$prob)
  propAMarked > propBMarked
}

speaker_test_fn_ <- function(df) {
  # Test one, check proportions
  filter(df, input=='a' & support == 'X a')$prob > filter(df, input=='b' & support == 'X b')$prob
}
```

### Corpus test
```{r corpus_helpers}
# runExperimentFn <- function(runFn, modelName, alpha, lambda, theta, n) {
#   return(function(expNum) {
#     rData <- newRData(modelName, alpha, lambda, theta, n)
#     df <- runFn(rData) %>%
#       mutate(modelName=modelName,
#              alpha=alpha,
#              lambda=lambda,
#              theta=theta,
#              n=n,
#              expNum=expNum)
#     df 
#   })
# }

createLM <- function(dfCorpus) {
  annotateUtterance <- function(x) {
    x <- gsub(' ', '', x)
    x <- paste0('^', x)
    x <- paste0(x, '$')
    return(x)
  }
  stream <- dfCorpus %>%
          mutate(annotatedUtterance=annotateUtterance(currUtterance)) %>%
          select(annotatedUtterance) %>%
          unlist %>%
          paste0(collapse='') %>%
          gsub(' ', '', ., fixed=TRUE)
  ng <- ngram::ngram(stream, n=2, sep='')  # bigram model
  dfNgramRaw <- data.frame(ngram::get.phrasetable(ng))
  dfNgramRaw$first <- sapply(dfNgramRaw$ngrams, function(x) {strsplit(x, '')}[[1]][1])
  dfNgramRaw$second <- sapply(dfNgramRaw$ngrams, function(x) {strsplit(x, '')}[[1]][3])
  dfNgramRaw <- dfNgramRaw %>%
    mutate(first=ifelse(first==' ', '*', first),
           second=ifelse(second==' ', '*', second))
  dfNgramRaw$theta <- unique(dfCorpus$theta)
  dfNgramRaw$alpha <- unique(dfCorpus$alpha)
  dfNgramRaw$lambda <- unique(dfCorpus$lambda)
  dfNgramRaw$expNum <- unique(dfCorpus$expNum)
  dfNgramRaw
}

run_corpus_analysis_ <- function(df_lm) {
  df_lm %>%
    group_by(second) %>%
    mutate(secondUncondProb=sum(prop),                # p(a|X) + p(a|^)
           lik=prop/secondUncondProb) %>%             # p(X|a) = p(X a) / p(a|X) + p(a|^)
    ungroup %>%
    mutate(logUncondProb=log2(secondUncondProb)) %>%  # p(a)
    ungroup %>%
    select(first, second, lik, logUncondProb) %>%
    filter(second %in% c(letters[1:5], '*'), first=='X') %>%
    group_by(second) %>%
    summarise(
      avgPostProb=mean(logUncondProb),
      avgLikelihood=mean(lik))
}

# Put in a function somewhere
# no_cores <- detectCores() - 1
# cl <- makeCluster(no_cores, type='FORK')
# registerDoParallel(cl)
# 
# simsExp1 <- foreach(i=seq(1, nIters), .packages=c('dplyr', 'rwebppl'), .combine=rbind) %dopar% runExp1(i)
# stopCluster(cl)
```

```{r test_3}
corpus_run_fn_ <- function(modelFile) {
  rData <- data.frame(modelName='S1', alpha=ALPHA, lambda=LAMBDA, theta=THETA, n=10000)
  df <- rwebppl::webppl(modelFile, data=rData, data_var='rData') %>%
    mutate(modelName=rData$modelName,
           alpha=rData$alpha,
           lambda=rData$lambda,
           theta=rData$theta,
           n=rData$n)
  # Parallelize runs below
  # -----------------------
  # no_cores <- detectCores() - 1
  # cl <- makeCluster(no_cores, type='FORK')
  # registerDoParallel(cl)
  # sims <- foreach(i=seq(1, nIters), .packages=c('dplyr', 'rwebppl'), .combine=rbind) %dopar% runExp1(i)
  df_lm <- createLM(df)
  df_summary <- run_corpus_analysis_(df_lm)
  df_summary
}

corpus_test_fn_ <- function(df) {
  return(
    (filter(df, second=='a')$avgPostProb < filter(df, second=='b')$avgPostProb) && 
    (filter(df, second=='a')$avgLikelihood > filter(df, second=='b')$avgLikelihood))
}
```

```{r run_tests_helper}
run_all_tests <- function(tests){
  sapply(tests, function(x) x())
}
```

```{r run_noisy_rsa1_tests}
noisy_rsa1_listener_file <- 'Models/testing/noisy-rsa1-listener.wppl'
noisy_rsa1_listener_test <- create_test(noisy_rsa1_listener_file, listener_run_fn_, listener_test_fn_)
noisy_rsa1_speaker_file <- 'Models/testing/noisy-rsa1-speaker.wppl'
noisy_rsa1_speaker_test <- create_test(noisy_rsa1_speaker_file, speaker_run_fn_, speaker_test_fn_)
noisy_rsa1_corpus_file <- 'Models/testing/noisy-rsa1-corpus.wppl'
noisy_rsa1_corpus_test <- create_test(noisy_rsa1_corpus_file, corpus_run_fn_, corpus_test_fn_)
noisy_rsa1_tests <- c(noisy_rsa1_listener_test, noisy_rsa1_speaker_test, noisy_rsa1_corpus_test)
run_all_tests(noisy_rsa1_tests)
```

```{r run_noisy_rsa1_tests}
noisy_rsa2_listener_file <- 'Models/testing/noisy-rsa2-listener.wppl'
noisy_rsa2_listener_test <- create_test(noisy_rsa2_listener_file, listener_run_fn_, listener_test_fn_)
noisy_rsa2_speaker_file <- 'Models/testing/noisy-rsa2-speaker.wppl'
noisy_rsa2_speaker_test <- create_test(noisy_rsa2_speaker_file, speaker_run_fn_, speaker_test_fn_)
noisy_rsa2_corpus_file <- 'Models/testing/noisy-rsa2-corpus.wppl'
noisy_rsa2_corpus_test <- create_test(noisy_rsa2_corpus_file, corpus_run_fn_, corpus_test_fn_)
noisy_rsa2_tests <- c(noisy_rsa2_listener_test, noisy_rsa2_speaker_test, noisy_rsa2_corpus_test)
run_all_tests(noisy_rsa2_tests)
```

## Basic RSA - no noise
```{r run_basic_rsa1_tests}
basic_rsa1_listener_file <- 'Models/testing/basic-rsa1-listener.wppl'
basic_rsa1_listener_test <- create_test(basic_rsa1_listener_file, listener_run_fn_, listener_test_fn_)
basic_rsa1_speaker_file <- 'Models/testing/basic-rsa1-speaker.wppl'
basic_rsa1_speaker_test <- create_test(basic_rsa1_speaker_file, speaker_run_fn_, speaker_test_fn_)
basic_rsa1_corpus_file <- 'Models/testing/basic-rsa1-corpus.wppl'
basic_rsa1_corpus_test <- create_test(basic_rsa1_corpus_file, corpus_run_fn_, corpus_test_fn_)
basic_rsa1_tests <- c(basic_rsa1_listener_test, basic_rsa1_speaker_test, basic_rsa1_corpus_test)
run_all_tests(basic_rsa1_tests)
```


