---
title: "Rsa-uid-model2"
author: "Ben"
date: "11/13/2017"
output: html_document
---

```{r packages, message=FALSE, warning=FALSE}
library(assertthat)
library(dplyr)
library(ggplot2)
library(rwebppl)
library(tidyr)
```

# Webppl modeling

## Get data from R env
```{r webpplModelStr1}
rEnvData <- "

// R Env data
var isListenerExp = function(rData) {
  var data = rData[0]
  return data['isListenerExp']
}

var getTheta = function(rData) {
  var data = rData[0]
  return data['theta']
}

var getInput = function(rData) {
  var data = rData[0]
  if (isListenerExp(rData)) {
    return data['input']
  }
  return {color:data['input']}
}

var getModelName = function(rData) {
  var data = rData[0]
  return data['modelName']
}

var inputStr = getInput(rData)
var THETA = getTheta(rData)

"
```

## RSA model variants
```{r webpplModelStr2}
rsa1 <- "

var deleteToken = function(utterance) {
  var tokens = utterance.split(' ')
  return remove(uniformDraw(tokens), tokens).join(' ')
}

// Alternative 
var replaceToken = function(utterance) {
  var tokens = utterance.split(' ')
  var index = randomInteger(tokens.length)
  return remove(uniformDraw(tokens), tokens).join(' ')
}

var noiseModel = function(utterance, theta) {
  return flip(theta) ? deleteToken(utterance) : utterance
}

var getColorFromUtterance = function(utterance) {
  return filter(function(x) {x != 'X'}, utterance.split(' ')).join(' ')
}

var getColorFromWorld = function(world) {
  return world['color']
}

var equalWorlds = function(w1, w2) {
  return w1['color'] == w2['color']
}

var meaning = function(utterance, world) {
  /* Meaning is truth-functional. */
  var utteranceColor = getColorFromUtterance(utterance)
  var worldColor = getColorFromWorld(world)
  return utteranceColor != 'X' | utteranceColor != '' ? utteranceColor == worldColor : false
}

var decode = function(utteranceProduced) {
  Infer({
    model() {
      var utteranceIntended = utterancePrior()
      var noisyUtterance = noiseModel(utteranceIntended, THETA)
      factor(utteranceProduced == noisyUtterance ? 0 : -Infinity)
      return utteranceIntended
    }
  })
}

/*****************
***************** RSA
*****************/
// var THETA = 0.95
var L0 = function(utterance) {
  Infer({
    model() {
      var world = worldPrior()
      var m = meaning(utterance, world)
      factor(m ? 0 : -Infinity)
      return world
    }
  })
}

var S1 = function(world) {
  Infer({
    model() {
      var utterance = utterancePrior()
      var L = L0(utterance)
      factor(L.score(world))
      return utterance
    }
  })
}

var _createListenerN = function(SN) {
  /* P(w|u) propto P(u|w)p(w) */
  return function(utterance) {
    Infer({
      model() {
        var world = worldPrior()  // P(w)
//         var intendedUtterance = sample(decode(utterance, THETA))
        var S = SN(world)  // P(u|w)
        factor(S.score(utterance))
        return world
      }
    })
  }
}

var _createSpeakerN = function(LN) {
  /* P(u|w) propto P(w|u)p(u) */
  return function(world) {
    Infer({
      model() {
        var utterance = utterancePrior()  // P(u)
//         var noisyUtterance = noiseModel(utterance, THETA)
        var L = LN(utterance)  // P(w|u)
        factor(L.score(world))
        return utterance
      }
    })
  }
}

var getModel = function(rData) {

  var L1 = _createListenerN(S1)
  var S2 = _createSpeakerN(L1)
  var L2 = _createListenerN(S2)
  var S3 = _createSpeakerN(L2)
  var L3 = _createListenerN(S3)
  var S4 = _createSpeakerN(L3)
  var modelName = getModelName(rData)
  if (modelName == 'L0') {
    return L0
  } else if (modelName == 'L1') {
    //console.log('returning L1 model')
    return L1
  } else if (modelName == 'L2') {
    //console.log('returning L2 model')
    return L2
  } else if (modelName == 'S1') {
    return S1
  } else if (modelName == 'S2') {
    return S2
  } else {
    return S3
  }
}

var model = getModel(rData)

// var markedUtt = 'X r'
// var unMarkedUtt = 'r'
// var w = {color:'r'}

model(inputStr)

"

rsa2 <- "

var deleteToken = function(utterance) {
  var tokens = utterance.split(' ')
  return remove(uniformDraw(tokens), tokens).join(' ')
}

// Alternative 
var replaceToken = function(utterance) {
  var tokens = utterance.split(' ')
  var index = randomInteger(tokens.length)
  return remove(uniformDraw(tokens), tokens).join(' ')
}

var noiseModel = function(utterance, theta) {
  return flip(theta) ? deleteToken(utterance) : utterance
}

var getColorFromUtterance = function(utterance) {
  return filter(function(x) {x != 'X'}, utterance.split(' ')).join(' ')
}

var getColorFromWorld = function(world) {
  return world['color']
}

var equalWorlds = function(w1, w2) {
  return w1['color'] == w2['color']
}

var meaning = function(utterance, world) {
  /* Meaning is truth-functional. */
  var utteranceColor = getColorFromUtterance(utterance)
  var worldColor = getColorFromWorld(world)
  return utteranceColor != 'X' | utteranceColor != '' ? utteranceColor == worldColor : false
}

var decode = function(utteranceProduced) {
  Infer({
    model() {
      var utteranceIntended = utterancePrior()
      var noisyUtterance = noiseModel(utteranceIntended, THETA)
      factor(utteranceProduced == noisyUtterance ? 0 : -Infinity)
      return utteranceIntended
    }
  })
}

/*****************
***************** RSA
*****************/
// var THETA = 0.95
var L0 = function(utterance) {
  Infer({
    model() {
      var world = worldPrior()
      var m = meaning(utterance, world)
      factor(m ? 0 : -Infinity)
      return world
    }
  })
}

var S1 = function(world) {
  Infer({
    model() {
      var utterance = utterancePrior()
      var noisyUtterance = noiseModel(utterance, THETA)
      var L = L0(noisyUtterance)
      factor(L.score(world))
      return utterance
    }
  })
}

var _createListenerN = function(SN) {
  /* P(w|u) propto P(u|w)p(w) */
  return function(utterance) {
    Infer({
      model() {
        var world = worldPrior()  // P(w)
//         var intendedUtterance = sample(decode(utterance, THETA))
        var S = SN(world)  // P(u|w)
        factor(S.score(utterance))
        return world
      }
    })
  }
}

var _createSpeakerN = function(LN) {
  /* P(u|w) propto P(w|u)p(u) */
  return function(world) {
    Infer({
      model() {
        var utterance = utterancePrior()  // P(u)
//         var noisyUtterance = noiseModel(utterance, THETA)
        var L = LN(utterance)  // P(w|u)
        factor(L.score(world))
        return utterance
      }
    })
  }
}

var getModel = function(rData) {

  var L1 = _createListenerN(S1)
  var S2 = _createSpeakerN(L1)
  var L2 = _createListenerN(S2)
  var S3 = _createSpeakerN(L2)
  var L3 = _createListenerN(S3)
  var S4 = _createSpeakerN(L3)
  var modelName = getModelName(rData)
  if (modelName == 'L0') {
    return L0
  } else if (modelName == 'L1') {
    //console.log('returning L1 model')
    return L1
  } else if (modelName == 'L2') {
    //console.log('returning L2 model')
    return L2
  } else if (modelName == 'S1') {
    return S1
  } else if (modelName == 'S2') {
    return S2
  } else {
    return S3
  }
}

var model = getModel(rData)

// var markedUtt = 'X r'
// var unMarkedUtt = 'r'
// var w = {color:'r'}

model(inputStr)

"

rsa3 <- "

var deleteToken = function(utterance) {
  var tokens = utterance.split(' ')
  return remove(uniformDraw(tokens), tokens).join(' ')
}

// Alternative 
var replaceToken = function(utterance) {
  var tokens = utterance.split(' ')
  var index = randomInteger(tokens.length)
  return remove(uniformDraw(tokens), tokens).join(' ')
}

var noiseModel = function(utterance, theta) {
  return flip(theta) ? deleteToken(utterance) : utterance
}

var getColorFromUtterance = function(utterance) {
  return filter(function(x) {x != 'X'}, utterance.split(' ')).join(' ')
}

var getColorFromWorld = function(world) {
  return world['color']
}

var equalWorlds = function(w1, w2) {
  return w1['color'] == w2['color']
}

var meaning = function(utterance, world) {
  /* Meaning is truth-functional. */
  var utteranceColor = getColorFromUtterance(utterance)
  var worldColor = getColorFromWorld(world)
  return utteranceColor != 'X' | utteranceColor != '' ? utteranceColor == worldColor : false
}

var decode = function(utteranceProduced) {
  Infer({
    model() {
      var utteranceIntended = utterancePrior()
      var noisyUtterance = noiseModel(utteranceIntended, THETA)
      factor(utteranceProduced == noisyUtterance ? 0 : -Infinity)
      return utteranceIntended
    }
  })
}

/*****************
***************** RSA
*****************/
// var THETA = 0.95
var L0 = function(utterance) {
  Infer({
    model() {
      var world = worldPrior()
      var intendedUtterance = sample(decode(utterance, THETA))
      var m = meaning(intendedUtterance, world)
      factor(m ? 0 : -Infinity)
      return world
    }
  })
}

var S1 = function(world) {
  Infer({
    model() {
      var utterance = utterancePrior()
      var noisyUtterance = noiseModel(utterance, THETA)
      var L = L0(noisyUtterance)
      factor(L.score(world))
      return utterance
    }
  })
}

var _createListenerN = function(SN) {
  /* P(w|u) propto P(u|w)p(w) */
  return function(utterance) {
    Infer({
      model() {
        var world = worldPrior()  // P(w)
//         var intendedUtterance = sample(decode(utterance, THETA))
        var S = SN(world)  // P(u|w)
        factor(S.score(utterance))
        return world
      }
    })
  }
}

var _createSpeakerN = function(LN) {
  /* P(u|w) propto P(w|u)p(u) */
  return function(world) {
    Infer({
      model() {
        var utterance = utterancePrior()  // P(u)
//         var noisyUtterance = noiseModel(utterance, THETA)
        var L = LN(utterance)  // P(w|u)
        factor(L.score(world))
        return utterance
      }
    })
  }
}

var getModel = function(rData) {

  var L1 = _createListenerN(S1)
  var S2 = _createSpeakerN(L1)
  var L2 = _createListenerN(S2)
  var S3 = _createSpeakerN(L2)
  var L3 = _createListenerN(S3)
  var S4 = _createSpeakerN(L3)
  var modelName = getModelName(rData)
  if (modelName == 'L0') {
    return L0
  } else if (modelName == 'L1') {
    //console.log('returning L1 model')
    return L1
  } else if (modelName == 'L2') {
    //console.log('returning L2 model')
    return L2
  } else if (modelName == 'S1') {
    return S1
  } else if (modelName == 'S2') {
    return S2
  } else {
    return S3
  }
}

var model = getModel(rData)

// var markedUtt = 'X r'
// var unMarkedUtt = 'r'
// var w = {color:'r'}

model(inputStr)

"
```

## Webbpl Worlds
```{r uniformWorldModelData}
allUniformData <- "

var colors = ['b', 'r']
var tokens = ['b', 'r', 'g', 'X']
var tokensDistr = Categorical({
  vs: tokens,
  ps: repeat(tokens.length, function() {return 1})
})
var tokensPrior = function() {
  return sample(tokensDistr)
}

var utterances = ['r', 'g', 'b', 'X r', 'X g', 'X b', 'X', '']
var utteranceDistr = Categorical({
  vs: utterances,
//  ps: [1, 1, 1, 0.50, 0.50, 0.50, 0.05, 0.05]
  ps: repeat(utterances.length, function() {return 1}) // Unif
})

var utterancePrior = function() {
  return sample(utteranceDistr)
}

var worlds = [{color:'r'}, {color:'b'}, {color:'g'}, {color: ''}]
var worldDistr = Categorical({
  vs: worlds,
//  ps: [0.5, 1, 1, 1]
  ps: repeat(worlds.length, function() {return 1}) // Unif
})

var worldPrior = function() {
  return sample(worldDistr)
}

"

utteranceCostWorldData <- "

var colors = ['b', 'r']
var tokens = ['b', 'r', 'g', 'X']
var tokensDistr = Categorical({
  vs: tokens,
  ps: repeat(tokens.length, function() {return 1})
})
var tokensPrior = function() {
  return sample(tokensDistr)
}

var utterances = ['r', 'g', 'b', 'X r', 'X g', 'X b', 'X', '']
var utteranceDistr = Categorical({
  vs: utterances,
  ps: [1, 1, 1, 0.50, 0.50, 0.50, 0.05, 0.05]
  //ps: repeat(utterances.length, function() {return 1}) // Unif
})

var utterancePrior = function() {
  return sample(utteranceDistr)
}

var worlds = [{color:'r'}, {color:'b'}, {color:'g'}, {color: ''}]
var worldDistr = Categorical({
  vs: worlds,
//  ps: [0.5, 1, 1, 1]
  ps: repeat(worlds.length, function() {return 1}) // Unif
})

var worldPrior = function() {
  return sample(worldDistr)
}

"

utteranceCostWorldRareData <- "

var colors = ['b', 'r']
var tokens = ['b', 'r', 'g', 'X']
var tokensDistr = Categorical({
  vs: tokens,
  ps: repeat(tokens.length, function() {return 1})
})
var tokensPrior = function() {
  return sample(tokensDistr)
}

var utterances = ['r', 'g', 'b', 'X r', 'X g', 'X b', 'X', '']
var utteranceDistr = Categorical({
  vs: utterances,
  ps: [1, 1, 1, 0.50, 0.50, 0.50, 0.05, 0.05]
  //ps: repeat(utterances.length, function() {return 1}) // Unif
})

var utterancePrior = function() {
  return sample(utteranceDistr)
}

var worlds = [{color:'r'}, {color:'b'}, {color:'g'}, {color: ''}]
var worldDistr = Categorical({
  vs: worlds,
  ps: [0.5, 1, 1, 0.5]
//  ps: repeat(worlds.length, function() {return 1}) // Unif
})

var worldPrior = function() {
  return sample(worldDistr)
}

"
```

### Combine webbpl model data
```{r combineWebpplPars}
combineWebpplModelEls <- function(...) {
  paste(...)
}
```

### Grid search
```{r gridSearchCode}
createGridSearchVectors <- function(dim1, dim2, dim3) {
  dim1_ <- rep(dim1, length(dim2) * length(dim3))
  dim2_ <- rep(rep(dim2, each=length(dim1)), length(dim3))
  dim3_ <- rep(dim3, each=length(dim1) * length(dim2))
  assertthat::assert_that(length(dim1_) == length(dim2_))
  assertthat::assert_that(length(dim2_) == length(dim3_))
  return(list(dim1_, dim2_, dim3_))
}
```

### Experiment helpers
```{r experimentCode}
runExperiment <- function(modelStr, modelName, input, theta, isListenerExp, expNum) {
  rData <- data.frame(modelName=modelName, input=input, theta=theta, isListenerExp=isListenerExp, stringsAsFactors=FALSE)
  df <- webppl(modelStr, data=rData, data_var='rData')
  df$modelName <- modelName
  df$input <- input
  df$theta <- theta
  df$expNum <- expNum
  df$isListenerExp <- isListenerExp
  return(df)
}
#runExperiment('L1', 'r', 0.99)

runExperiments <- function(modelStr, modelNames, inputs, thetas, isListenerExp) {
  expNums <- seq(1, length(thetas))
  data <- mapply(function(a, b, c, d, e, f) {
    runExperiment(a, b, c, d, e, f)
  }, modelStr, modelNames, inputs, thetas, isListenerExp, expNums, SIMPLIFY=FALSE)
  #data <- parallel::mcmapply(function(a, b, c, d, e, f) {
  #  runExperiment(a, b, c, d, e, f)
  #}, modelStr, modelNames, inputs, thetas, isListenerExp, expNums, SIMPLIFY=FALSE, mc.cores=4)
  df <- do.call(rbind, data)
  rownames(df) <- 1:nrow(df)
  return(df)
}
#modelNames <- c('L0', 'L0', 'L1', 'L1')
#inputs <- c('X r', 'r', 'X r', 'X r')
#thetas <- rep(0.5, 4)
#runExperiments(modelNames, inputs, thetas)
```


# All `listener` combinations
```{r webpplModelStr}
webpplModelStr1 <- combineWebpplModelEls(rEnvData, allUniformData, rsa1)
webpplModelStr2 <- combineWebpplModelEls(rEnvData, allUniformData, rsa2)
webpplModelStr3 <- combineWebpplModelEls(rEnvData, allUniformData, rsa3)
webpplModelStr4 <- combineWebpplModelEls(rEnvData, utteranceCostWorldData, rsa1)
webpplModelStr5 <- combineWebpplModelEls(rEnvData, utteranceCostWorldData, rsa2)
webpplModelStr6 <- combineWebpplModelEls(rEnvData, utteranceCostWorldData, rsa3)
webpplModelStr7 <- combineWebpplModelEls(rEnvData, utteranceCostWorldRareData, rsa1)
webpplModelStr8 <- combineWebpplModelEls(rEnvData, utteranceCostWorldRareData, rsa2)
webpplModelStr9 <- combineWebpplModelEls(rEnvData, utteranceCostWorldRareData, rsa3)
```

```{r runListenerExperimentsSetup}
thetas_ <- seq(0.05, 0.95, 0.10)
#rep(thetas_, length(models_) * length(inputs_))
inputs_ <- c('X r', 'r')
#rep(rep(inputs_, each=length(thetas_)), length(models_))
models_ <- c('L0', 'L1', 'L2')
#rep(models_, each=length(thetas_)*length(inputs_))

gridSearchVecs <- createGridSearchVectors(thetas_, inputs_, models_)
thetas <- gridSearchVecs[[1]]
inputs <- gridSearchVecs[[2]]
models <- gridSearchVecs[[3]]
expType <- rep(TRUE, length(models))
```

```{r runExperiments}
# Webppl strs
webpplModelStrs1 <- rep(webpplModelStr1, length(models))
webpplModelStrs2 <- rep(webpplModelStr2, length(models))
webpplModelStrs3 <- rep(webpplModelStr3, length(models))
webpplModelStrs4 <- rep(webpplModelStr4, length(models))
webpplModelStrs5 <- rep(webpplModelStr5, length(models))
webpplModelStrs6 <- rep(webpplModelStr6, length(models))
webpplModelStrs7 <- rep(webpplModelStr7, length(models))
webpplModelStrs8 <- rep(webpplModelStr8, length(models))
webpplModelStrs9 <- rep(webpplModelStr9, length(models))

# Run models
dfListener1 <- runExperiments(webpplModelStrs1, models, inputs, thetas, expType) %>%
  mutate(expType='1')
dfListener2 <- runExperiments(webpplModelStrs2, models, inputs, thetas, expType) %>%
  mutate(expType='2')
dfListener3 <- runExperiments(webpplModelStrs3, models, inputs, thetas, expType) %>%
  mutate(expType='3')
dfListener4 <- runExperiments(webpplModelStrs4, models, inputs, thetas, expType) %>%
  mutate(expType='4')
dfListener5 <- runExperiments(webpplModelStrs5, models, inputs, thetas, expType) %>%
  mutate(expType='5')
dfListener6 <- runExperiments(webpplModelStrs6, models, inputs, thetas, expType) %>%
  mutate(expType='6')
dfListener7 <- runExperiments(webpplModelStrs7, models, inputs, thetas, expType) %>%
  mutate(expType='7')
dfListener8 <- runExperiments(webpplModelStrs8, models, inputs, thetas, expType) %>%
  mutate(expType='8')
dfListener9 <- runExperiments(webpplModelStrs9, models, inputs, thetas, expType) %>%
  mutate(expType='9')

# Agg
dfListener <- rbind(dfListener1,
                    dfListener2,
                    dfListener3,
                    dfListener4,
                    dfListener5,
                    dfListener6,
                    dfListener7,
                    dfListener8,
                    dfListener9)
```

```{r listenerPlot1}
dfListener %>%
  filter(modelName %in% c('L1', 'L2')) %>%
  filter(input %in% c('X r', 'r')) %>%
  filter(color %in% c('r', '')) %>%
  mutate(meaning=ifelse(color=='r', 'correct', ifelse(color=='', 'incorrect_nothing', 'incorrect_other'))) %>%
  filter(meaning == 'correct') %>%
  ggplot(aes(x=theta, y=prob, col=input)) +
    geom_line(aes(lty=meaning), size=1.25) +
    ggtitle("UID Listener models") +
    xlab("Noise amount THETA") +
    ylab("p(m|u)") +
    facet_grid(modelName~expType) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))
```


```{r}
dfListener %>%
  filter(modelName %in% c('L1')) %>%
  filter(input %in% c('X r', 'r')) %>%
  filter(color %in% c('r', '')) %>%
  mutate(meaning=ifelse(color=='r', 'correct', ifelse(color=='', 'incorrect_nothing', 'incorrect_other'))) %>%
  filter(meaning == 'correct' & expType=='6') %>%
  ggplot(aes(x=theta, y=prob, col=input)) +
    geom_line(aes(lty=meaning), size=1.25) +
    ggtitle("UID L1 Listener") +
    xlab("Noise amount THETA") +
    ylab("p(m|u)") +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))
```


# All `Speaker` data
```{r}
thetas_ <- seq(0.05, 0.95, 0.10)
#rep(thetas_, length(models_) * length(inputs_))
inputs_ <- c('r')
#rep(rep(inputs_, each=length(thetas_)), length(models_))
models_ <- c('S1', 'S2', 'S3')
#rep(models_, each=length(thetas_)*length(inputs_))

gridSearchVecs <- createGridSearchVectors(thetas_, inputs_, models_)
thetas <- gridSearchVecs[[1]]
inputs <- gridSearchVecs[[2]]
models <- gridSearchVecs[[3]]
expType <- rep(FALSE, length(models))
expType <- rep(FALSE, length(models))
```

```{r}
# Webppl strs
webpplModelStrs1 <- rep(webpplModelStr1, length(models))
webpplModelStrs2 <- rep(webpplModelStr2, length(models))
webpplModelStrs3 <- rep(webpplModelStr3, length(models))
webpplModelStrs4 <- rep(webpplModelStr4, length(models))
webpplModelStrs5 <- rep(webpplModelStr5, length(models))
webpplModelStrs6 <- rep(webpplModelStr6, length(models))
webpplModelStrs7 <- rep(webpplModelStr7, length(models))
webpplModelStrs8 <- rep(webpplModelStr8, length(models))
webpplModelStrs9 <- rep(webpplModelStr9, length(models))

# Run models
dfSpeaker1 <- runExperiments(webpplModelStrs1, models, inputs, thetas, expType) %>%
  mutate(expType='1')
dfSpeaker2 <- runExperiments(webpplModelStrs2, models, inputs, thetas, expType) %>%
  mutate(expType='2')
dfSpeaker3 <- runExperiments(webpplModelStrs3, models, inputs, thetas, expType) %>%
  mutate(expType='3')
dfSpeaker4 <- runExperiments(webpplModelStrs4, models, inputs, thetas, expType) %>%
  mutate(expType='4')
dfSpeaker5 <- runExperiments(webpplModelStrs5, models, inputs, thetas, expType) %>%
  mutate(expType='5')
dfSpeaker6 <- runExperiments(webpplModelStrs6, models, inputs, thetas, expType) %>%
  mutate(expType='6')
dfSpeaker7 <- runExperiments(webpplModelStrs7, models, inputs, thetas, expType) %>%
  mutate(expType='7')
dfSpeaker8 <- runExperiments(webpplModelStrs8, models, inputs, thetas, expType) %>%
  mutate(expType='8')
dfSpeaker9 <- runExperiments(webpplModelStrs9, models, inputs, thetas, expType) %>%
  mutate(expType='9')

# Combine
dfSpeaker <- rbind(dfSpeaker1,
                   dfSpeaker2,
                   dfSpeaker3,
                   dfSpeaker4,
                   dfSpeaker5,
                   dfSpeaker6,
                   dfSpeaker7,
                   dfSpeaker8,
                   dfSpeaker9)
```

```{r}
dfSpeaker %>%
  filter(support %in% c('X r', 'r')) %>%
  ggplot(aes(x=theta, y=prob, col=support)) +
    geom_line(size=1.25) +
    facet_grid(modelName~expType) +
    ggtitle("UID speaker models") +
    ylab('p(u|m)')+
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))
```


```{r}
dfSpeaker %>%
  filter(support %in% c('X r', 'r')) %>%
  filter(modelName=='S1' & expType=='5') %>%
  ggplot(aes(x=theta, y=prob, col=support)) +
    geom_line(size=1.25) +
    ggtitle("S1 Speaker data") +
    xlab("Noise amount THETA") +
    ylab("p(u|m)") +
    ylim(0, 1) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))
```



# Model 1 - Uniform utterance cost

Utterances:
`ps: repeat(utterances.length, function() {return 1}) // Unif`

## Listener data

```{r listenerPlot2}
dfListener1 %>% 
  mutate(input=ifelse(input=='X r', 'Xr', input)) %>%
  select(color, modelName, theta, input, prob) %>%
  spread(input, prob) %>%
  mutate(probDiff=Xr-r) %>%
  filter(color %in% c('r', '')) %>%
  filter(modelName %in% c('L2')) %>%
  ggplot(aes(x=theta, y=probDiff, fill=color)) +
    geom_bar(stat='identity', position='dodge') +
    ylab("Prefer marked version") +
    xlab("Noise (theta)") +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))
```

## Speaker data
```{r runSpeakerExperimentsSetup}
thetas_ <- seq(0.05, 0.95, 0.45)
#rep(thetas_, length(models_) * length(inputs_))
inputs_ <- c('r')
#rep(rep(inputs_, each=length(thetas_)), length(models_))
models_ <- c('S1', 'S2')
#rep(models_, each=length(thetas_)*length(inputs_))

gridSearchVecs <- createGridSearchVectors(thetas_, inputs_, models_)
thetas <- gridSearchVecs[[1]]
inputs <- gridSearchVecs[[2]]
models <- gridSearchVecs[[3]]
expType <- rep(FALSE, length(models))
webpplModelStr <- combineModels(rEnvData, uniformWorldData, rsaModel)
webpplModelStrs <- rep(webpplModelStr1, length(models))
expType <- rep(FALSE, length(models))
```

```{r runExperiments}
dfSpeaker1 <- runExperiments(webpplModelStrs, models, inputs, thetas, expType)
```

```{r speakerDataProcess}
dfSpeaker1 %>%
  filter(support %in% c('X r', 'r', '')) %>%
  filter(modelName %in% c('S1', 'S2')) %>%
  ggplot(aes(x=theta, y=prob, col=support)) +
    geom_line() +
    facet_wrap(~modelName)
```


# Model 2 - Utterance costs

`var utterances = ['r', 'g', 'b', 'X r', 'X g', 'X b', 'X', '']`

## Listener
```{r utteranceCostWorldData}

```

```{r listenerModelStr}
webpplModelStr <- combineModels(rEnvData, utteranceCostWorldData, rsaModel)
```

```{r }
thetas_ <- seq(0.05, 0.95, 0.05)
#rep(thetas_, length(models_) * length(inputs_))
inputs_ <- c('X r', 'r')
#rep(rep(inputs_, each=length(thetas_)), length(models_))
models_ <- c('L0', 'L1', 'L2')
#rep(models_, each=length(thetas_)*length(inputs_))

gridSearchVecs <- createGridSearchVectors(thetas_, inputs_, models_)
thetas <- gridSearchVecs[[1]]
inputs <- gridSearchVecs[[2]]
models <- gridSearchVecs[[3]]
expType <- rep(TRUE, length(models))
webpplModelStrs <- rep(webpplModelStr, length(models))
```

```{r}
dfListener2 <- runExperiments(webpplModelStrs, models, inputs, thetas, expType)
```

```{r listenerPreprocess}
dfListener2 <- dfListener2 %>%
  mutate(utteranceType=ifelse('X' == strsplit(input, ' ')[[1]][1], 'marked', 'unmarked')) %>%
  mutate(meaning=ifelse(color=='r', '_red', ifelse(color=='', 'empty', 'other')))
```

```{r listenerPlot3}
dfListener2 %>%
  filter(modelName %in% c('L1')) %>%
  filter(color %in% c('r', '')) %>%
  ggplot(aes(x=theta, y=prob, col=input)) +
    geom_line(aes(lty=meaning)) +
    ggtitle("Listener data") +
    xlab("Noise amount THETA") +
    ylab("p(m|u)") +
    facet_wrap(~modelName) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))
```

```{r tempListenerPlotsForPreso, eval=FALSE}
dfListener2 %>%
  filter(modelName %in% c('L1')) %>%
  filter(color %in% c('r', '')) %>%
  filter(meaning=='_red' & input=='r') %>%
  ggplot(aes(x=theta, y=prob, col=input)) +
    geom_line(aes(lty=meaning), size=1.25) +
    ggtitle("L1 Listener data") +
    xlab("Noise amount THETA") +
    ylab("p(m|u)") +
    ylim(0, 1) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))

dfListener2 %>%
  filter(modelName %in% c('L1')) %>%
  filter(color %in% c('r', '')) %>%
  filter(input=='r') %>%
  ggplot(aes(x=theta, y=prob, col=input)) +
    geom_line(aes(lty=meaning), size=1.25) +
    ggtitle("L1 Listener data") +
    xlab("Noise amount THETA") +
    ylab("p(m|u)") +
    ylim(0, 1) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))

dfListener2 %>%
  filter(modelName %in% c('L1')) %>%
  filter(color %in% c('r', '')) %>%
  filter(meaning %in% c('_red', 'other')) %>%
  ggplot(aes(x=theta, y=prob, col=input)) +
    geom_line(aes(lty=meaning), size=1.25) +
    ggtitle("L1 Listener data") +
    xlab("Noise amount THETA") +
    ylab("p(m|u)") +
    ylim(0, 1) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))

dfListener2 %>%
  filter(modelName %in% c('L1')) %>%
  filter(color %in% c('r', '')) %>%
  ggplot(aes(x=theta, y=prob, col=input)) +
    geom_line(aes(lty=meaning), size=1.25) +
    ggtitle("L1 Listener data") +
    xlab("Noise amount THETA") +
    ylab("p(m|u)") +
    ylim(0, 1) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))

```


```{r listenerPlot4}
dfListener2 %>% 
  mutate(input=ifelse(input=='X r', 'Xr', input)) %>%
  select(color, modelName, theta, input, prob) %>%
  spread(input, prob) %>%
  mutate(probDiff=Xr-r) %>%
  filter(color %in% c('r', '')) %>%
  filter(modelName %in% c('L1', 'L2')) %>%
  ggplot(aes(x=theta, y=probDiff, fill=color)) +
    geom_bar(stat='identity', position='dodge') +
    ylab("Prefer marked version") +
    xlab("Noise (theta)") +
    facet_wrap(~modelName) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))
```

## Speaker

```{r runSpeakerExperimentsSetup2}
createGridSearchVectors <- function(dim1, dim2, dim3) {
  dim1_ <- rep(dim1, length(dim2) * length(dim3))
  dim2_ <- rep(rep(dim2, each=length(dim1)), length(dim3))
  dim3_ <- rep(dim3, each=length(dim1) * length(dim2))
  assertthat::assert_that(length(dim1_) == length(dim2_))
  assertthat::assert_that(length(dim2_) == length(dim3_))
  return(list(dim1_, dim2_, dim3_))
}

thetas_ <- seq(0.05, 0.95, 0.05)
#rep(thetas_, length(models_) * length(inputs_))
inputs_ <- c('r')
#rep(rep(inputs_, each=length(thetas_)), length(models_))
models_ <- c('S1', 'S2', 'S3')
#rep(models_, each=length(thetas_)*length(inputs_))

gridSearchVecs <- createGridSearchVectors(thetas_, inputs_, models_)
thetas <- gridSearchVecs[[1]]
inputs <- gridSearchVecs[[2]]
models <- gridSearchVecs[[3]]
expType <- rep(FALSE, length(models))
webpplModelStr <- combineModels(rEnvData, utteranceCostWorldData, rsaModel)
webpplModelStrs <- rep(webpplModelStr, length(models))
expType <- rep(FALSE, length(models))
```

```{r runExperiments2}
dfSpeaker2 <- runExperiments(webpplModelStrs, models, inputs, thetas, expType)
```

```{r speakerDataProcess2}
dfSpeaker2 %>%
  filter(support %in% c('X r', 'r')) %>%
  filter(modelName %in% c('S2')) %>%
  ggplot(aes(x=theta, y=prob, col=support)) +
    geom_line(size=1.25) +
    facet_wrap(~modelName) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))
```


```{r tempSpeakerPlotForPreso}
dfSpeaker %>%
  filter(support %in% c('r')) %>%
  filter(modelName %in% c('S2')) %>%
  ggplot(aes(x=theta, y=prob, col=support)) +
    geom_line(size=1.25) +
    ggtitle("S2 Speaker data") +
    xlab("Noise amount THETA") +
    ylab("p(u|m)") +
    ylim(0, 1) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))


dfSpeaker %>%
  filter(support %in% c('X r', 'r')) %>%
  filter(modelName %in% c('S2')) %>%
  ggplot(aes(x=theta, y=prob, col=support)) +
    geom_line(size=1.25) +
    ggtitle("S2 Speaker data") +
    xlab("Noise amount THETA") +
    ylab("p(u|m)") +
    ylim(0, 1) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))

dfSpeaker %>%
  filter(modelName %in% c('S2')) %>%
  ggplot(aes(x=theta, y=prob, col=support)) +
    geom_line(size=1.25) +
    ggtitle("S2 Speaker data") +
    xlab("Noise amount THETA") +
    ylab("p(u|m)") +
    ylim(0, 1) +
    theme_bw() +
    theme(axis.text.x=element_text(angle=90, hjust=1), 
          plot.title=element_text(hjust=0.5, size=14))
```

