---
title: "rsa-uid-discourse-20180227"
author: "Ben"
date: "2/27/2018"
output: html_document
---

# Contents:

(1) Bootstrapped entropy decomposition comparing `contextAwareS2` to `baseline`
(2) Bootstrapped entropy decomposition with `contextAwareS2` varying alphas

```{r libraries, message=FALSE, warning=FALSE}
rm(list = ls())  # clear workspace

library(doParallel)
library(dplyr)
library(entropy)
library(gganimate)
library(ggplot2)
library(ggthemes)
library(tidyr)
library(zoo)

source("run-helpers.R")
```

```{r getModel}
fPath <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/Models/discourse/rsa-uid-discourse-20180215.wppl'
m <- getModelFile(fPath)
```

```{r runFn}
run <- createRunFn(m)
runFn <- function(i, targetDistr, nUtterances, resultType, alpha) {
  dTemp <- data.frame(targetDistr=targetDistr, nUtterances=nUtterances, resultType=resultType, alpha=alpha)
  df <- run(dTemp) %>%
    mutate(runNum=i, targetDistr=targetDistr, alpha=alpha, resultType=resultType, nUtterances=nUtterances)
  df
}
# test_rnFn <- runFn(1, 'T1', 100, 'contextAwareS2', 10, FALSE)
```

# (1) Bootstrapped entropy decomposition comparing `contextAwareS2` to `baseline`
```{r contextAwareS2-vs-baseline-alpha20-compare}
nUtterances <- 120
binSize <- 20
df10_contextAware <- runParVaryingAlpha(runFn,
                                       targetDistr='T1',
                                       resultType='contextAwareS2',
                                       alpha=10, 
                                       nUtterances=nUtterances, 
                                       binSize=binSize)

df10_baseline <- runParVaryingAlpha(runFn,
                                   targetDistr='T1', 
                                   resultType='baseline',
                                   alpha=10, 
                                   nUtterances=nUtterances, 
                                   binSize=binSize)
```



```{r preprocess-contextAwareS2-vs-baseline-alpha20-compare}
df10_join <- rbind(df10_contextAware, df10_baseline)
```


Check proportions
```{r plot-proportions-contextAwareS2-vs-baseline-alpha20-compare}
df10_join %>%
  filter(runNum %in% seq(1, 4)) %>%
  ggplot(aes(x=binVal, y=prop, fill=utterance)) +
    # geom_vline(data=df_agg %>% select(resultType, posBin) %>% unique, aes(xintercept=posBin), lty=2) +
    geom_bar(stat='identity', alpha=0.5) +
    facet_grid(~resultType) +
    theme_few() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_wrap(resultType~runNum, nrow=2)
```

Check entropy
```{r plot-unconditional-entropy-contextAwareS2-vs-baseline-alpha20-compare}
df10_join %>%
  group_by(resultType, runNum, binVal) %>%
  summarise(entropy=entropy::entropy(norm_(n), unit=c('log2'))) %>%
  ggplot(aes(x=binVal, y=entropy)) +
    # geom_jitter(alpha=0.2, width=0.05, height=0) +
    geom_smooth(method='loess') +
    theme_few() +
    facet_grid(~resultType)
```

Populate entropy data...
```{r run-hy-factorization-contextAwareS2-vs-basline-compare}
breaks <- seq(0, nUtterances, by=binSize)
end <- length(breaks)-1
endPoints <- seq(1, end)
topicProps <- c(0.6, 0.25, 0.15, 0.15, 0.1, 0.1)

df10_join_bs <- df10_join %>%
  group_by(resultType, runNum) %>%
  do(getEntropyData(., endPoints, topicProps))
```

Plot.
```{r plot-hy-factorization-contextAwareS2-vs-basline-compare}
df10_join_bs %>%
  gather(type,val, c(uncondEntropy, MI, hY)) %>%
  ggplot(aes(x=endPoint, y=val, col=type)) +
    # geom_jitter(alpha=0.2, width=0.25) +
    geom_smooth(method='loess', level=0.99) +
    theme_few() +
    facet_grid(~resultType)
```

(2) Bootstrapped entropy decomposition with `contextAwareS2` varying alphas
```{r run-bootstrapped-contextAware-varying-alpha}
# Settings
nUtterances <- 120
binSize <- 20

# Run alphas
alphas <- c(2, 4, 8, 16, 32)
df_agg <- data.frame()
for (alpha_ in alphas) {
  cat("\nrunning alpha: ", alpha_, " ")
  curr_df <- runParVaryingAlpha(runFn, targetDistr='T1', alpha=alpha_, nUtterances=nUtterances, binSize=binSize)
  df_agg <- rbind(df_agg, curr_df)
}
```

Populate entropy data...
```{r run-entropy-bootstrapped-contextAware-varying-alpha}
breaks <- seq(0, nUtterances, by=binSize)
end <- length(breaks)-1
endPoints <- seq(1, end)
topicProps <- c(0.6, 0.25, 0.15, 0.15, 0.1, 0.1)

df_agg_bs <- df_agg %>%
  group_by(alpha, runNum) %>%
  do(getEntropyData(., endPoints, topicProps))
```

Plot.
```{r plot-entropy-bootstrapped-contextAware-varying-alpha}
df_agg_bs %>%
  gather(type,val, c(uncondEntropy, MI, hY)) %>%
    ggplot(aes(x=endPoint, y=val, col=type)) +
      # geom_jitter(alpha=0.2, width=0.25) +
      geom_smooth(method='loess', level=0.99) +
      theme_few() +
      facet_wrap(~alpha)
```
