---
title: "rsa-uid-discourse-abstractions-20180304"
author: "Ben"
date: "3/4/2018"
output: html_document
---

# Contents:

(1) Single run discourse abstractions model.
(2) Bootstrapped simulations.


```{r libraries, message=FALSE, warning=FALSE}
rm(list = ls())  # clear workspace

library(doParallel)
library(dplyr)
library(entropy)
library(gganimate)
library(ggplot2)
library(ggthemes)
library(tidyr)
library(zoo)

source("run-helpers.R")
```

Load model.
```{r getModel}
fPath <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/Models/discourse/rsa-uid-discourse-abstractions-conjugate-updates-20180314.wppl'
m <- getModelFile(fPath)
```

Setup.
```{r runFn}
run <- createRunFn(m)
runFn <- function(i, targetDistr, nUtterances, resultType, alpha) {
  dTemp <- data.frame(targetDistr=targetDistr, nUtterances=nUtterances, resultType=resultType, alpha=alpha)
  df <- run(dTemp) %>%
    mutate(runNum=i, targetDistr=targetDistr, alpha=alpha, resultType=resultType)
  df
}
```

# (1) Single run models

Single model runs -- contrast `contexUnawareS2` and `contextAwareS2`.
```{r singleRun}
ALPHA <- 6
TOPIC <- 'T1'
NUM_UTTERANCES <- 200
# TODO (BP): This needs to be fixed...
# df_baseline <- runFn(1, 
#                      targetDistr=TOPIC, 
#                      nUtterances=NUM_UTTERANCES, 
#                      resultType='baseline', 
#                      alpha=ALPHA)
df_contextUnawareS2 <- runFn(1,
                            targetDistr=TOPIC,
                            nUtterances=NUM_UTTERANCES,
                            resultType='contextUnawareS2',
                            alpha=ALPHA)
df_contextAwareS2 <- runFn(1, 
                           targetDistr=TOPIC, 
                           nUtterances=NUM_UTTERANCES, 
                           resultType='contextAwareS2', 
                           alpha=ALPHA)

df_agg <- rbind(df_contextUnawareS2, df_contextAwareS2)
# df_agg <- rbind(df_baseline, df_contextUnwareS2, df_contextAwareS2)
```


S2 speaker more likely to use ambiguous utterance `x` when listener is confident in the (correct) topic.
```{r}
df_agg %>%
  mutate(xHere=ifelse(utterance == 'x', 1, 0),
         yHere=ifelse(utterance == 'y', 0.5, 0)) %>%
  gather(utteranceLoc, val, c(xHere, yHere)) %>%
  ggplot(aes(x=utteranceNum, y=val)) +
    geom_jitter(alpha=0.50, height=0.1, width=0, aes(col=utterance), size=2) +
    geom_line(aes(x=utteranceNum, y=T1), col='blue') +
    geom_line(aes(x=utteranceNum, y=T2), col='red') +
    geom_line(aes(x=utteranceNum, y=T3), col='green') +
    geom_line(aes(x=utteranceNum, y=T4), col='purple') +
    theme_few() +
    facet_grid(~resultType)
```


# (2) Bootstrapped simulations.

Run sims.

Note: This simulation is now quite slow running taking around 60min to run `100` sims of `75` utterances.
```{r run-sims}
nSims <- 100

# Parallelization setup
no_cores <- detectCores() - 1
cl <- makeCluster(no_cores, type='FORK')
registerDoParallel(cl)

# Run sims
ptm <- proc.time()
sims <- foreach(i=seq(1, nSims), .packages=c('dplyr', 'rwebppl'), .combine=rbind) %dopar% 
  runFn(i, targetDistr='T1', nUtterances=NUM_UTTERANCES, resultType='contextAwareS2', alpha=ALPHA)
stopCluster(cl)
etm <- proc.time() - ptm
cat("runtime: ", etm[3] / 60)

# df_par <- runParVaryingAlpha(runFn=runFn,
#                              targetDistr='T1',
#                              resultType='contextAwareS2',
#                              nUtterances=30,
#                              nSims=25,
#                              alpha=ALPHA,
#                              binSize=1)
```

Plot utterances along listener posterior probabilyt of (correct) topic. We expect more red (and yellow) points in later positions.
```{r plot-sims}
sims %>%
  mutate(`P(Topic|Utterances)`=ifelse(utterance=='x', 0.5, ifelse(utterance=='y', 0.75, 0)),
         utterance=ifelse(utterance=='x', '[a, d]', ifelse(utterance=='y', '[b, c]', utterance))) %>%
  ggplot(aes(x=utteranceNum, y=`P(Topic|Utterances)`)) +
    geom_smooth(method='loess', aes(y=T1)) +
    geom_jitter(aes(col=utterance), width=0.5, height=0.1, alpha=0.25, size=2) +
    ylab("Listener posterior probability of target topic") +
    xlab("Utterance position") +
    ggtitle("S2 production (points) and listener topic inference (line)\nn=100 simulations") +
    theme_few() +
    theme(plot.title=element_text(hjust = 0.5))
```

Empirical probability of ambiguous utterance as a function of utterance position.
```{r}
sims %>%
  mutate(isX=utterance=='x',
         isA=utterance=='a',
         isB=utterance=='b') %>%
  group_by(utteranceNum) %>%
  summarise(n=n(),
            numX=sum(isX),
            probX=numX/n,
            stdX=sqrt(probX*(1-probX)/n),
            numA=sum(isA),
            probA=numA/n,
            stdA=sqrt(probA*(1-probA)/n),
            numB=sum(isB),
            probB=numB/n,
            stdB=sqrt(probB*(1-probB)/n),
            avgT1=mean(T1),
            avgT2=mean(T2),
            avgT3=mean(T3),
            avgT4=mean(T4)) %>%
  # Check relations to other utterances here:
  # gather(type, val, c(probX, probA, probB)) %>%
  # ggplot(aes(x=utteranceNum, y=val, col=type)) +
  ggplot(aes(x=utteranceNum, y=probX)) +
    # geom_smooth(aes(y=avgT1), col='red', alpha=0.2) +
    # geom_smooth(aes(y=avgT2), col='green', alpha=0.2) +
    geom_jitter(size=2, alpha=0.75, width=2, height=0) +
    geom_smooth(method='lm', alpha=0.5, col='blue') +
    geom_errorbar(aes(ymin=probX-2*stdX, ymax=probX+2*stdX), alpha=0.5, width=0) +
    ylab("Empirical probability") +
    xlab("Utterance position") +
    ggtitle("Empirical probability S2 uses ambiguous utterance\nn=100 simulations") +
    theme_few() +
    theme(plot.title=element_text(hjust = 0.5))
```

Check proportions
```{r plot-proportions-contextAwareS2-vs-baseline-alpha20-compare}
binSize=10
df_binned <- addUtteranceBins(sims, nUtterances=NUM_UTTERANCES, binSize=binSize)
df_filled <- fillUtteranceProportions(df_binned, binSize)
```


```{r}
df_filled %>% 
  filter(runNum %in% seq(1, 5)) %>%
  ggplot(aes(x=binVal, y=prop, fill=utterance)) +
    # geom_vline(data=df_agg %>% select(resultType, posBin) %>% unique, aes(xintercept=posBin), lty=2) +
    geom_bar(stat='identity', alpha=0.5) +
    facet_grid(~resultType) +
    theme_few() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_wrap(~runNum)
```


```{r}
df_filled %>%
  group_by(resultType, runNum, binVal) %>%
  summarise(entropy=entropy::entropy(norm_(n), unit=c('log2'))) %>%
  ggplot(aes(x=binVal, y=entropy)) +
    geom_smooth(method='lm') +
    # geom_smooth(method='lm') +
    # geom_jitter(alpha=0.2, width=0.05, height=0) +
    theme_few()
```

