---
title: "Rsa-uid-plots-for-cocolab"
author: "Ben"
date: "1/27/2018"
output: html_document
---

```{r warning=FALSE, message=FALSE}
library(directlabels)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(tidyr)

source("run-helpers.R")
setwd('/Users/benpeloquin/Desktop/Projects/rsa_uid/')
```

```{r}
basic_rsa1_listener_file <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/Models/testing/basic-rsa1-listener.wppl'
basic_rsa1_speaker_file <- 'Models/testing/basic-rsa1-speaker.wppl'
basic_rsa1_corpus_file <- 'Models/testing/basic-rsa1-corpus.wppl'

noisy_rsa1_listener_file <- 'Models/testing/noisy-rsa1-listener.wppl'
noisy_rsa1_speaker_file <- 'Models/testing/noisy-rsa1-speaker.wppl'
noisy_rsa1_corpus_file <- 'Models/testing/noisy-rsa1-corpus.wppl'
```

# Speakers

```{r basic-vs-noisy-speaker-data}
inputs <- c('a', 'b', 'c', 'd')
dirPath <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/'
d_s1_store <- data.frame()
for (i in inputs) {
  df_brsa_speaker <- speaker_run_fn(paste0(dirPath, basic_rsa1_speaker_file), modelName='S5', alpha=5, speakerInput=i, theta=0.3) %>%
    mutate(model='basic_rsa',
           input=i)
  df_nrsa_speaker <- speaker_run_fn(paste0(dirPath, noisy_rsa1_speaker_file), modelName='S5', alpha=5, speakerInput=i, theta=0.3) %>%
    mutate(model='noisy_rsa',
           input=i)
  d_s1_store <- rbind(d_s1_store, df_brsa_speaker, df_nrsa_speaker)  
}

surprisals <- data.frame(d=c('a', 'b', 'c', 'd'), 
                         vals=c(0.2, 0.5, 1, 2.5)) %>%
  mutate(normed_vals=vals/sum(vals),
         surprisal=-log2(normed_vals))
d_s1_store <- left_join(d_s1_store, surprisals, by=c('input'='d'))
```

```{r basic-vs-noisy-speaker-plot}
d_s1_store %>%
  mutate(referent=ifelse(support %in% c('X a', 'a'), 'a', 
                          ifelse(support %in% c('X b', 'b'), 'b',
                                 ifelse(support %in% c('X c', 'c'), 'c', 'd'))),
         marked=grepl('X', support)) %>%
  filter(referent==input, marked) %>%
  ggplot(aes(x=paste0(referent, ' (', round(surprisal, 2), ')'), y=prob)) +
    geom_bar(aes(fill=model), position='dodge', stat='identity') +
    ylab("P( world | UID-cue )") +
    xlab("world") +
    theme_few()
```

```{r}
surprisals <- data.frame(d=c('a', 'b', 'c', 'd'), 
                         vals=c(0.2, 0.5, 1, 2.5)) %>%
  mutate(normed_vals=vals/sum(vals),
         surprisal=-log2(normed_vals))

dir_path <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/'
p <- paste0(dir_path, noisy_rsa1_speaker_file)
inputs <- c('a', 'b', 'c', 'd')
thetas <- seq(0.1, 0.9, by=0.1)
d_s1_store <- data.frame()
for (theta in thetas) {
  cat("theta:", theta, "\n")
  for (i in inputs) {
    cat("input:", i, " ")
    curr_df <- speaker_run_fn(p, modelName='S5', alpha=5, speakerInput=i, theta=theta) %>%
      mutate(speakerInput=i,
             theta=theta)
    d_s1_store <- rbind(d_s1_store, curr_df)
  }
}

df_s1 <- d_s1_store %>%
  mutate(d=speakerInput)
df_s1 <- left_join(df_s1, surprisals, by='d') %>%
  mutate(surprisal=round(surprisal, digits=2))
```

Single theta example
```{r speaker-plot1}
df_s1 %>%
  mutate(trueWorld=ifelse(support %in% c('X a', 'a'), 'a', 
                          ifelse(support %in% c('X b', 'b'), 'b',
                                 ifelse(support %in% c('X c', 'c'), 'c', 'd'))),
         marked=grepl('X', support)) %>%
  filter(round(theta, 1)==0.3) %>%
  filter(grepl('X', support), trueWorld==speakerInput) %>%
  ggplot(aes(x=as.factor(surprisal), y=prob, fill=support)) +
    geom_bar(stat='identity', position='dodge') +
    xlab("object surprisal log2(P(w)") +
    ylab("Likelihood of UID cue") +
    ylim(0, 1) +
    scale_color_ptol(aes(support)) +
    theme_few() +
    theme(legend.position="none")
```

```{r speaker-plot2}
df_s1 %>%
  mutate(trueWorld=ifelse(support %in% c('X a', 'a'), 'a', 
                          ifelse(support %in% c('X b', 'b'), 'b',
                                 ifelse(support %in% c('X c', 'c'), 'c', 'd'))),
         marked=grepl('X', support)) %>%
  filter(grepl('X', support), trueWorld==speakerInput) %>%
  ggplot(aes(x=theta, y=prob, col=trueWorld)) +
    geom_line() +
    xlab("Theta") +
    ylab("Likelihood of UID cue") +
    geom_dl(aes(label=trueWorld), method = list(dl.combine("last.points"), cex=2)) +
    theme_few() +
    theme(legend.position="none")
```

# Listeners

### Basic RSA - listener plot
```{r}
df_l1 <- listener_run_fn(basic_rsa1_listener_file)

df_l1 %>%
  ggplot(aes(x=d, y=prob)) +
    geom_bar(stat='identity') +
    xlab("world") +
    ylab("p(world|UID-cue)") +
    theme_few()
```


## Noisy RSA - listener plot
```{r nrsa-lplot1-data}
dir_path <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/'
p <- paste0(dir_path, noisy_rsa1_listener_file)
df_l2 <- listener_run_fn(p, theta=0.95)

surprisals <- data.frame(d=c('a', 'b', 'c', 'd'), 
                         vals=c(0.2, 0.5, 1, 2.5)) %>%
  mutate(normed_vals=vals/sum(vals),
         surprisal=round(-log2(normed_vals), 2))

df_l2 <- left_join(df_l2, surprisals, by='d')
```

Note I'm no longer using this in the langcog presentation. Instead using the next plot in `combine-brsa1-nrsa1-plot` chunk.
```{r nrsa-lplot1}
df_l2 %>%
  ggplot(aes(x=as.factor(surprisal), y=prob, fill=d)) +
    geom_bar(stat='identity') +
    geom_text(aes(x=as.factor(surprisal), y=prob+0.05, ymax=prob, label=d), 
              position = position_dodge(width=1), size=6, col='black') +
    xlab("object surprisal log2(P(w)") +
    ylab("P(world | UID-cue )") +
    ylim(0, 1) +
    theme_few() +
    theme(legend.position="none")
```

```{r combine brsa1-nrsa1-preprocess}
df_listener_join <- left_join(df_l2, df_l1, by='d') %>%
  rename(noisy_rsa=prob.x,
         basic_rsa=prob.y)
```

```{r combine-brsa1-nrsa1-plot}
df_listener_join %>%
  gather(model, prob, c(noisy_rsa, basic_rsa)) %>%
  ggplot(aes(x=as.factor(paste(d, '(', surprisal, ')')), y=prob)) +
    # geom_line(aes(col=model), group=2) +
    geom_bar(stat='identity', position='dodge', aes(fill=model)) +
    # geom_text(aes(x=as.factor(surprisal), y=prob+0.05, ymax=prob, label=d), 
    #           position = position_dodge(width=1), size=6, col='black') +
    xlab("object surprisal log2(P(w)") +
    ylab("P(world | UID-cue )") +
    ylim(0, 1) +
    theme_few()
```


```{r nrsa1-vary-theta-data}
p <- paste0(dir_path, noisy_rsa1_listener_file)
inputs <- c('a', 'b', 'c', 'd', 'X a', 'X b', 'X c', 'X d', 'X')
thetas <- seq(0.1, 0.9, by=0.1)
df_listeners <- data.frame()
for (i in inputs) {
  cat("\ninput:", i, ' ')
  for (theta in thetas) {
    cat("theta:", theta, ' ')
    curr_df <- listener_run_fn(p, theta=theta, listenerInput=i) %>%
      mutate(input=i,
             theta=theta)
    df_listeners <- rbind(df_listeners, curr_df)
  }
}
```
```{r nrsa-plot2-varying-theta}
# df_listeners %>%
#   ggplot(aes(x=d, y=prob)) +
#     geom_bar(stat='identity', position='dodge') +
#     facet_grid(theta~input)

df_listeners %>%
  filter(input=='X') %>%
  rename(world=d) %>%
  ggplot(aes(x=theta, y=prob, col=world)) +
    geom_line(size=1.5, alpha=0.8) +
    xlab("Theta") +
    ylab("P( world | UID-cue )") +
    ylim(0, 1) +
    theme_few()

```

Given multiple referent, varying in surprisal, the implicature to the high(er) surprisal items is a function of noise.
```{r nrsa-lplot2}
# d %>%
#   ggplot(aes(x=d, y=prob, col=theta)) +
#     geom_line(aes(group=theta)) +
#     scale_colour_ptol(aes(theta)) +
#     xlab("world") +
#     ylab("p( world | UID-cue )") +
#     theme_few() 

# d %>%
#   mutate(surprisal_=as.factor(paste(d,"(", as.character(surprisal), ")"))) %>%
#   mutate(theta=as.numeric(theta)) %>%
#   ggplot(aes(x=theta, y=prob, fill=d)) +
#     geom_bar(stat='identity') +
#     facet_wrap(~surprisal_) +
#     theme_few() +
#     theme(legend.position='none')

# d %>%
#   mutate(surprisal_=as.character(surprisal)) %>%
#   mutate(theta=as.numeric(theta)) %>%
#   ggplot(aes(x=theta, y=prob, col=d)) +
#     geom_line(aes(group=surprisal_), size=1.5, alpha=0.8) +
#     geom_dl(aes(label=d), method = list(dl.combine("last.points"), cex=2)) +
#     xlab("Theta") +
#     ylab("P( world | UID-cue )") +
#     ylim(0, 1) +
#     theme_few() +
#     theme(legend.position="none")
```

# Corpus plots
```{r corpus-brsa-vs-nrsa-data}
dir_path <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/'
n_corpus <- paste0(dir_path, noisy_rsa1_corpus_file)
b_corpus <- paste0(dir_path, basic_rsa1_corpus_file)
# function(modelFile, modelName, alpha, lambda, theta, nUtterances)
df_nrsa_corpus <- corpus_run_fn(n_corpus, "S3", 5, 1, 0.3, 1000) %>%
  mutate(model='noisy_rsa')
df_brsa_corpus <- corpus_run_fn(b_corpus, "S3", 5, 1, 0.3, 1000)  %>%
  mutate(model='basic_rsa')
df_corpsus_comb <- rbind(df_nrsa_corpus, df_brsa_corpus)
```

```{r corpus-brsa-vs-nrsa-plot}
df_corpsus_comb %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=model)) +
      geom_line() +
      geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
      geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
      xlab("log( P( u1 | u2 ))") +
      ylab("Likelihood of full form") +
      ylim(0, 1) +
      theme_few()
```


```{r corpus-nrsa-vary-theta-data}
dir_path <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/'
n_corpus <- paste0(dir_path, noisy_rsa1_corpus_file)
thetas <- seq(0.1, 0.9, by=0.1)
df_corpus2 <- data.frame()
for (theta in thetas) {
  cat("Running theta:", theta, " ")
  curr_df <- corpus_run_fn(n_corpus, modelName="S5", theta=theta, alpha=5, nUtterances=1000) %>%
    mutate(modelName="S5",
           theta=theta)
  df_corpus2 <- rbind(df_corpus2, curr_df)
}
```

```{r corpus-nrsa-vary-theta-plot}
df_corpus2 %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=as.factor(theta))) +
    geom_line() +
    theme_few()
```


### `CAUTION` - slow running

Notice cached data below...
```{r corpus-nrsa-vary-theta-and-model-data}
dir_path <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/'
p_corpus <- paste0(dir_path, noisy_rsa1_corpus_file)

thetas <- seq(0.1, 0.9, by=0.1)
models <- c('S1', 'S2', 'S3', 'S4', 'S5')
df_corpuses <- data.frame()
for (model in models ) {
  for (theta in thetas) {
    cat("Running theta:", theta, ' model:', model, '\n')
    df_curr_run <- corpus_run_fn(p_corpus, modelName=model, alpha=5, lambda=1, theta=theta, nUtterances=500) %>%
      mutate(theta=theta,
             modelName=model)
    df_corpuses <- rbind(df_corpuses, df_curr_run)
  }
}

# Store and read from cache
cache_path <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/Data/cache-corpus-run-vary-speaker-theta-20180208.csv'
# write.csv(df_corpuses, file=cache_path)
```

```{r corpus-nrsa-vary-theta-and-model-plot2}
df_corpuses %>%
  mutate(noise=as.factor(theta)) %>%
  filter(noise==0.3) %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood)) +
    geom_line(aes(col=modelName)) +
    # geom_ribbon(aes(ymin=ymin,ymax=ymax, fill=modelName), alpha=0.3) +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
    xlab("log( P( u1 | u2 ))") +
    ylab("Likelihood of full form") +
    ylim(0, 1) +
    theme_few()
```



```{r corpus-nrsa-vary-theta-and-model-plot2}
df_corpuses %>%
  mutate(noise=as.factor(theta))%>%
  # filter(modelName=='S5') %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood)) +
    geom_line(aes(col=modelName)) +
    # geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
    # geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
    xlab("log( P( u1 | u2 ))") +
    ylab("Likelihood of full form") +
    ylim(0, 1) +
    theme_few() +
    # theme(legend.position='none') +
    facet_wrap(~noise)

# d_noisy_rsa1 %>%
#   filter(theta==0.3, modelName=='S5') %>%
#   ggplot(aes(x=avgPostProb, y=avgLikelihood)) +
#     geom_line() +
#     geom_errorbar(aes(ymax=ymax, ymin=ymin, width=0)) +
#     geom_errorbarh(aes(xmax=xmax, xmin=xmin, height=0)) +
#     xlab("log( P( u1 | u2 ))") +
#     ylab("Likelihood of full form") +
#     ylim(0, 1) +
#     theme_few()
```


