---
title: "rsa-uid-discourse-noise-ambiguity-model-20180322"
author: "Ben"
date: "3/22/2018"
output: html_document
---

```{r libraries, message=FALSE, warning=FALSE}
rm(list = ls())  # clear workspace

library(doParallel)
library(dplyr)
library(entropy)
library(gganimate)
library(ggplot2)
library(ggthemes)
library(tidyr)
library(zoo)
setwd("/Users/benpeloquin/Desktop/Projects/rsa_uid/")
source("Analyses/run-helpers.R")
```

Load model.
```{r getModel}
fPath <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/Models/discourse/rsa-uid-discourse-noise-ambiguity-20180322.wppl'
m <- getModelFile(fPath)
```

Setup.
```{r runFn}
run <- createRunFn(m)
runFn <- function(i, targetDistr, nUtterances, resultType, alpha, theta) {
  dTemp <- data.frame(targetDistr=targetDistr, nUtterances=nUtterances, resultType=resultType, alpha=alpha, theta=theta)
  df <- run(dTemp) %>%
    mutate(runNum=i, targetDistr=targetDistr, alpha=alpha, theta=theta, resultType=resultType)
  df
}
```

```{r singleRun}
ALPHA <- 8
THETA <- 0.15
TOPIC <- 'T1'
NUM_UTTERANCES <- 80
# TODO (BP): This needs to be fixed...
df_baseline <- runFn(1,
                     targetDistr=TOPIC,
                     nUtterances=NUM_UTTERANCES,
                     resultType='baseline',
                     alpha=ALPHA,
                     theta=THETA)
df_contextUnawareS2 <- runFn(1,
                            targetDistr=TOPIC,
                            nUtterances=NUM_UTTERANCES,
                            resultType='contextUnawareS4',
                            alpha=ALPHA,
                            theta=THETA)
df_contextAwareS2 <- runFn(1, 
                           targetDistr=TOPIC, 
                           nUtterances=NUM_UTTERANCES, 
                           resultType='contextAwareS4', 
                           alpha=ALPHA,
                           theta=THETA)

df_agg <- rbind(df_contextUnawareS2, df_contextAwareS2, df_baseline)
# df_agg <- rbind(df_baseline, df_contextUnwareS2, df_contextAwareS2)
```

S2 speaker more likely to use ambiguous utterance `x` when listener is confident in the (correct) topic.
```{r}
lvls <- as.factor(df_agg$utterance)

df_agg %>%
  rowwise() %>%
  mutate(utteranceYPos=which(utterance==levels(lvls))/10) %>%
  # mutate(xHere=ifelse(utterance == 'x' | utterance == 'X x', 1, 0),
  #        yHere=ifelse(utterance == 'y' | utterance == 'X y', 0.5, 0)) %>%
  # gather(utteranceLoc, val, c(xHere, yHere)) %>%
  ggplot(aes(x=utteranceNum, y=utteranceYPos, col=utterance)) +
    # geom_jitter(alpha=0.50, height=0.1, width=0, aes(col=utterance), size=2) +
    geom_point() +
    geom_line(aes(x=utteranceNum, y=T1), col='blue') +
    geom_line(aes(x=utteranceNum, y=T2), col='red') +
    geom_line(aes(x=utteranceNum, y=T3), col='green') +
    geom_line(aes(x=utteranceNum, y=T4), col='purple') +
    theme_few() +
    facet_grid(~resultType)
```

# (2) Bootstrapped simulations.

Run sims. Note (n=500 sims at nUttt=200 utterances takes 30min to run)

```{r run-sims}
nSims <- 100

# Parallelization setup
no_cores <- detectCores() - 1
cl <- makeCluster(no_cores, type='FORK')
registerDoParallel(cl)

# Run sims
ptm <- proc.time()
sims <- foreach(i=seq(1, nSims), .packages=c('dplyr', 'rwebppl'), .combine=rbind) %dopar% 
  runFn(i, targetDistr='T1', nUtterances=NUM_UTTERANCES, resultType='contextAwareS4', alpha=ALPHA, theta=THETA)
stopCluster(cl)
etm <- proc.time() - ptm
cat("runtime: ", etm[3] / 60)
```
Plot utterances along listener posterior probability of (correct) topic. We expect more red (and yellow) points in later positions.
```{r plot-sims}
sims %>%
  mutate(`P(Topic|Utterances)`=ifelse(utterance=='x' | utterance=='X x', 0.5, 
                                      ifelse(utterance=='y' | utterance=='X y', 0.75, 0)),
         utterance=ifelse(utterance=='x', '[a, d]', 
                          ifelse(utterance=='X x', 'X[a, d]', 
                                 ifelse(utterance=='y', '[b, c]', ifelse(utterance=='X y', 'X[b, c]', utterance))))) %>%
  ggplot(aes(x=utteranceNum, y=`P(Topic|Utterances)`)) +
    geom_smooth(method='loess', aes(y=T1)) +
    geom_jitter(aes(col=utterance), width=0.5, height=0.1, alpha=0.25, size=2) +
    ylab("Listener posterior probability of target topic") +
    xlab("Utterance position") +
    ggtitle(paste("S4 production (points) and listener topic inference (line)\nn=",nSims," simulations")) +
    theme_few() +
    theme(plot.title=element_text(hjust = 0.5))
```

Empirical probability of ambiguous utterance as a function of utterance position.
```{r}
sims %>%
  mutate(isAmbiguous=(utterance=='x' | utterance=='y' | utterance=='X x' | utterance=='X y'),
         isX=utterance=='x',
         isA=utterance=='a',
         isB=utterance=='b') %>%
  group_by(utteranceNum) %>%
  summarise(n=n(),
            numAmb=sum(isAmbiguous),
            probAmb=numAmb/n,
            stdAmb=sqrt(probAmb*(1-probAmb)/n),
            numX=sum(isX),
            probX=numX/n,
            stdX=sqrt(probX*(1-probX)/n),
            numA=sum(isA),
            probA=numA/n,
            stdA=sqrt(probA*(1-probA)/n),
            numB=sum(isB),
            probB=numB/n,
            stdB=sqrt(probB*(1-probB)/n),
            avgT1=mean(T1),
            avgT2=mean(T2),
            avgT3=mean(T3),
            avgT4=mean(T4)) %>%
  # Check relations to other utterances here:
  # gather(type, val, c(probX, probA, probB)) %>%
  # ggplot(aes(x=utteranceNum, y=val, col=type)) +
  ggplot(aes(x=utteranceNum, y=probAmb)) +
    # geom_smooth(aes(y=avgT1), col='red', alpha=0.2) +
    # geom_smooth(aes(y=avgT2), col='green', alpha=0.2) +
    geom_point(size=2, alpha=0.75) +
    geom_smooth(method='loess', alpha=0.5, col='blue') +
    # geom_smooth(method = "lm", formula = y ~ poly(x, 2)) +
    geom_errorbar(aes(ymin=probAmb-2*stdAmb, ymax=probAmb+2*stdAmb), alpha=0.5, width=0) +
    ylab("Empirical probability") +
    xlab("Utterance position") +
    ggtitle(paste("Empirical probability S4 uses ambiguous utterance\nn=",nSims," simulations")) +
    theme_few() +
    theme(plot.title=element_text(hjust = 0.5))
```

Check proportions
```{r plot-proportions-contextAwareS2-vs-baseline-alpha20-compare}
binSize=10
df_binned <- addUtteranceBins(sims, nUtterances=NUM_UTTERANCES, binSize=binSize)
df_filled <- fillUtteranceProportions(df_binned, binSize)
```


```{r}
df_filled %>% 
  filter(runNum %in% seq(1, 5)) %>%
  ggplot(aes(x=binVal, y=prop, fill=utterance)) +
    # geom_vline(data=df_agg %>% select(resultType, posBin) %>% unique, aes(xintercept=posBin), lty=2) +
    geom_bar(stat='identity', alpha=0.5) +
    facet_grid(~resultType) +
    theme_few() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_wrap(~runNum)
```

## Genzel & Charniak (2002) effect
```{r}
df_filled %>%
  group_by(resultType, runNum, binVal) %>%
  summarise(entropy=entropy::entropy(norm_(n), unit=c('log2'))) %>%
  ggplot(aes(x=binVal, y=entropy)) +
    # geom_smoothmethod='lm') +
    geom_smooth(method='lm', col='blue') +
    # geom_jitter(alpha=0.1, width=0.05, height=0) +
    ggtitle("Uncond entropy (Genzel & Charniak) effect") +
    ylab("Unconditional entropy (of bin)") +
    xlab("Binned utterance position") +
    theme_few() +
    theme(plot.title = element_text(hjust = 0.5))
```


```{r}
run_corpus_analysis2 <- function(df_lm, ci=0.90) {
  df_lm %>%
    group_by(expNum, second) %>%
    mutate(secondTotalProp=sum(prop),       # p(a|X) + p(a|^)
           lik=prop/secondTotalProp) %>%    # p(X|a) = p(X a) / p(a|X) + p(a|^)
    ungroup %>%
    group_by(expNum) %>%
    mutate(postProb=log2(secondTotalProp/sum(prop))) %>%  # p(a)
    ungroup %>%
    select(expNum, first, second, lik, postProb) %>%
    filter(second %in% c(letters[1:5], 'x', 'y'), first=='X') %>%
    group_by(second) %>%
    summarise(
      avgPostProb=median(postProb),
      # sdPostProb=sd(postProb),
      # xmin=avgPostProb-2*sdPostProb,
      # xmax=avgPostProb+2*sdPostProb,
      xmin=quantile(postProb, probs=c(1-ci)),
      xmax=quantile(postProb, probs=c(ci)),
      avgLikelihood=median(lik),
      # sdLikelihood=sd(lik),
      # ymin=avgLikelihood-2*sdLikelihood,
      # ymax=avgLikelihood+2*sdLikelihood,
      ymin=quantile(lik, probs=c(1-ci)),
      ymax=quantile(lik, probs=c(ci)),
      n=n())
}
```

# Levy & Jaeger (2007) effect
```{r process-sims}
# Early utterances
sims_processed_early <- sims %>%
  filter(utteranceNum < 20) %>%
  mutate(currUtterance=utterance,
         expNum=runNum) %>%
  process_sims(.) %>%
  run_corpus_analysis(., ci=0.90) %>%
  mutate(simType='early')

# Middle utterances
sims_processed_middle <- sims %>%
  filter(utteranceNum >= 20 & utteranceNum < 60) %>%
  mutate(currUtterance=utterance,
         expNum=runNum) %>%
  process_sims(.) %>%
  run_corpus_analysis(., ci=0.90) %>%
  mutate(simType='middle')

# Late utterances
sims_processed_late <- sims %>%
  filter(utteranceNum >= 60) %>%
  mutate(currUtterance=utterance,
         expNum=runNum) %>%
  process_sims(.) %>%
  run_corpus_analysis(., ci=0.90) %>%
  mutate(simType='late')


# All utterances
sims_full <- sims %>%
  mutate(currUtterance=utterance,
         expNum=runNum) %>%
  process_sims(.) %>%
  run_corpus_analysis(., ci=0.90) %>%
  mutate(simType='full')

sims_all <- rbind(sims_processed_early, sims_processed_middle, sims_processed_late, sims_full)
sims_all$simType <- factor(sims_all$simType, 
                           levels=c("early", "middle", "late", "full"), 
                           labels=c("<20", "20 <= x < 60", ">60", "all"))
```


```{r}
sims_full %>%
  filter(second %in% letters[1:4]) %>%
  # mutate(alph=ifelse(second %in% letters[1:4], 1, 0.8)) %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=as.factor(second))) +
    geom_point(size=2.5) +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, alpha=0.2), width=0) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin, alpha=0.2), height=0) +
    ylab('Likelihood of marked version') +
    xlab('lm surprisal -- mean(log(p(utt))') +
    ggtitle("UID effect in discourse model") +
    theme_few() +
    theme(plot.title = element_text(hjust = 0.5))
```


```{r}
sims_all %>% 
  # filter(second %in% letters[1:4]) %>%
  # mutate(alph=ifelse(second %in% letters[1:4], 1, 0.8)) %>%
  ggplot(aes(x=avgPostProb, y=avgLikelihood, col=as.factor(second))) +
    geom_point(size=2.5) +
    geom_errorbar(aes(ymax=ymax, ymin=ymin, alpha=0.2), width=0) +
    geom_errorbarh(aes(xmax=xmax, xmin=xmin, alpha=0.2), height=0) +
    ylab('Likelihood of marked version') +
    xlab('mean(log(p(utt))') +
    ylim(0, 1) +
    theme_few() +
    facet_wrap(~simType, nrow=1)
```

# Single speaker behavior
```{r single-speaker-model-file}
dPath <- '/Users/benpeloquin/Desktop/Projects/rsa_uid/Models/discourse/'
fPath <- 'rsa-uid-discourse-noise-ambiguity-single-run-20180322.wppl'
p <- paste0(dPath, fPath)
m <- getModelFile(p)
```

```{r single-speaker-run-fn}
run <- createRunFn(m)
runFn <- function(i, world, targetDistr, alpha, theta, data) {
  
  # Topic posterior
  dTemp <- data.frame(world=world,
                      targetDistr=targetDistr,
                      alpha=alpha, 
                      resultType='topicPosterior',
                      theta=theta,
                      data=data)
  df_topic_post <- run(dTemp) %>%
    mutate(runNum=i,
           world=world,
           targetDistr=targetDistr, 
           alpha=alpha, 
           theta=theta,
           data=data) %>%
    select(support, prob, data) %>%
    rename(topicPost=prob)
    
  # Speaker data
  dTemp <- data.frame(world=world,
                      targetDistr=targetDistr,
                      resultType='speaker',
                      alpha=alpha, 
                      theta=theta,
                      data=data)
  df_speaker <- run(dTemp) %>%
    mutate(runNum=i,
           world=world,
           targetDistr=targetDistr, 
           alpha=alpha, 
           theta=theta,
           data=data)
  
  df_merge <- merge(df_speaker, df_topic_post, by='data')
  df_merge
}
```

```{r single-speaker-test}
ALPHA <- 6
THETA <- 0.15
TOPIC <- 'T1'
# DATA <- paste0(rep('a', 10), collapse=' ')
DATA <- ''

d <- runFn(1, 'a', TOPIC, ALPHA, THETA, DATA)
```


```{r single-speaker-run}
# data
worlds <- c('a', 'b', 'c', 'd')
datas <- c('', paste0(rep('a', 1), collapse=' '), paste0(rep('a', 10), collapse=' '))
thetas <- seq(0.1, 0.9, by=0.1)

df_agg <- data.frame()
for (w in worlds) {
  cat(paste('\nworld:', w))
  for (d in datas) {
    for (theta in thetas) {
      
      currDf <- runFn(1, w, TOPIC, ALPHA, theta, d) %>%
        mutate(world=w,
               data=d)
      df_agg <- rbind(df_agg, currDf)
    }
  }
}
```

```{r single-speaker-plot1}
getDataAmount <- function(data) {
  nchar(gsub(' ', '' , data))
}

df_agg %>%
  filter(support=='T1') %>%
  filter((world=='a' & (utterance == 'X a' | utterance == 'a')) | 
           (world=='b' & (utterance == 'X b' | utterance == 'b')) |
           (world=='c' & (utterance == 'X c' | utterance == 'c')) |
           (world=='d' & (utterance == 'X d' | utterance == 'd'))) %>%
  rowwise %>%
  mutate(utteranceLen=getDataAmount(utterance),
         numPrevUtterances=getDataAmount(data),
         marked=ifelse(utteranceLen==2, 'marked', 'unmarked')) %>%
  select(prob, world, theta, numPrevUtterances, marked) %>%
  spread(marked, prob) %>%
  group_by(world, numPrevUtterances) %>%
  mutate(highestMarked=max(marked),
         lowestUnmarked=min(unmarked)) %>%
  gather(type, prob, c(marked, unmarked)) %>%
  # mutate(cross=ifelse(marked > unmarked, theta, 0)) %>%
  ggplot(aes(x=theta, y=prob, lty=type)) +
    geom_line() +
    theme_few() +
    ylab('Speaker posterior of unmarked or marked utterance') +
    xlab('Noise level (theta)') +
    ggtitle('Speaker likelihood of marking across thetas') +
    facet_grid(numPrevUtterances~world) +
    geom_line(aes(y=highestMarked), col='blue', alpha=0.2) +
    geom_line(aes(y=lowestUnmarked), col='red', alpha=0.2) +
    theme(plot.title=element_text(hjust = 0.5))
```


```{r single-speaker-plot2}
getDataAmount <- function(data) {
  nchar(gsub(' ', '' , data))
}


df_agg %>%
  filter(support=='T1') %>%
  filter((world=='a' & (utterance == 'X a' | utterance == 'a')) | 
           (world=='b' & (utterance == 'X b' | utterance == 'b')) |
           (world=='c' & (utterance == 'X c' | utterance == 'c')) |
           (world=='d' & (utterance == 'X d' | utterance == 'd'))) %>%
  rowwise %>%
  mutate(utteranceLen=getDataAmount(utterance),
         numPrevUtterances=getDataAmount(data),
         marked=ifelse(utteranceLen==2, 'marked', 'unmarked')) %>%
  select(prob, world, theta, numPrevUtterances, marked) %>%
  spread(marked, prob) %>%
  mutate(diff=marked-unmarked) %>%
  # mutate(cross=ifelse(marked > unmarked, theta, 0)) %>%
  ggplot(aes(x=theta, y=diff, fill=world)) +
    geom_area(alpha=0.4) +
    theme_few() +
    ylab('Speaker posterior of unmarked or marked utterance') +
    xlab('Noise level (theta)') +
    ggtitle('Speaker likelihood of marking across thetas') +
    facet_grid(world~numPrevUtterances)
```

